### 2014-04-22 paper.tex
removed useless stuff

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE! 
%% User assumes all risk.
%% In no event shall IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%
%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
%%*************************************************************************

\documentclass[10pt,        % Don't change the font size!
               a4paper,     % Don't change the paper size!
               journal,     % Journal paper format
%               draft       % Enable this parameter to get a draft version.
               ]{IEEEtran}
\makeatletter

\def\markboth#1#2{\def\leftmark{\@IEEEcompsoconly{\sffamily}\MakeUppercase{\protect#1}}%
\def\rightmark{\@IEEEcompsoconly{\sffamily}\MakeUppercase{\protect#2}}}
\makeatother

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% *** CITATION PACKAGES ***
\usepackage{cite}
\usepackage[pdftex]{graphicx}
\graphicspath{{../png/}}
\DeclareGraphicsExtensions{.png}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{algorithmic}
\usepackage{array}
\usepackage{mdwmath}
\usepackage{mdwtab}
\usepackage{eqparbox}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{stfloats}
\usepackage{url}
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Seminar Topic}
%
%
% author names and IEEE memberships
% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
% a structure at a ~ so this keeps an author's name from being broken across
% two lines.
% use \thanks{} to gain access to the first footnote area
% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
% was not built to handle multiple paragraphs
%

\author{John~Smith}% <-this % stops a space
%\thanks{M. Shell is with the Department
%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

% note the % following the last \IEEEmembership and also \thanks - 
% these prevent an unwanted space from occurring between the last author name
% and the end of the author line. i.e., if you had this:
% 
% \author{....lastname \thanks{...} \thanks{...} }
%                     ^------------^------------^----Do not want these spaces!
%
% a space would be appended to the last name and could cause every name on that
% line to be shifted left slightly. This is one of those "LaTeX things". For
% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
% "AB" then you have to do: "\textbf{A}\textbf{B}"
% \thanks is no different in this regard, so shield the last } of each \thanks
% that ends a line with a % and do not let a space in before the next \thanks.
% Spaces after \IEEEmembership other than the last one are OK (and needed) as
% you are supposed to have spaces between the names. For what it is worth,
% this is a minor point as most people would not even notice if the said evil
% space somehow managed to creep in.

% The paper headers
%\markboth{Hauptseminar Sicherheit in der Informationstechnik, Sommersemester 2011}%
\markboth{Advanced Seminar for Security in Information Technology, Summer Term 2011}%
{John Smith: Seminar Topic}

%\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

\begin{abstract}
%\boldmath
Put the abstract here. The abstract should be 200-300 words long and must
answer the following questions:
\begin{itemize}
   \item Why is this topic important and interesting?
   \item What is the topic?
\end{itemize}
It's a good writing style to answer the "why" question first by putting the
topic in a broader context.

\end{abstract}
% IEEEtran.cls defaults to using nonbold math in the Abstract.
% This preserves the distinction between vectors and scalars. However,
% if the journal you are submitting to favors bold math in the abstract,
% then you can use LaTeX's standard command \boldmath at the very start
% of the abstract to achieve this. Many IEEE journals frown on math
% in the abstract anyway.

% Note that keywords are not normally used for peerreview papers.
%\begin{IEEEkeywords}
%IEEEtran, journal, \LaTeX, paper, template.
%\end{IEEEkeywords}

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
% \IEEEpeerreviewmaketitle

%\section{Einleitung}
\section{Introduction}
In general a scientific paper starts with an introduction to the topic.

%\section{Stand der Technik}
\section{State of the Art}
In this section an analysis of the available literature on the topic is done.
This section may be split or subdivided into several sections or subsections.

\subsection{Subsection Heading Here}
Subsection text here.

% needed in second column of first page if using \IEEEpubid
%\IEEEpubidadjcol

\subsubsection{Subsubsection Heading Here}
Subsubsection text here.

\section{Please Note}
Regular research papers need at least two additional sections here. One section
for contributions and methods and one section for the results. For seminar
papers these sections can be omitted.

% An example of a floating figure using the graphicx package.
% Note that \label must occur AFTER (or within) \caption.
% For figures, \caption should occur after the \includegraphics.
% Note that IEEEtran v1.7 and later has special internal code that
% is designed to preserve the operation of \label within \caption
% even when the captionsoff option is in effect. However, because
% of issues like this, it may be the safest practice to put all your
% \label just after \caption rather than within \caption{}.
%
% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
% option should be used if it is desired that the figures are to be
% displayed while in draft mode.
%
%\begin{figure}[!t]
%\centering
%\includegraphics[width=2.5in]{myfigure}
% where an .eps filename suffix will be assumed under latex, 
% and a .pdf suffix will be assumed for pdflatex; or what has been declared
% via \DeclareGraphicsExtensions.
%\caption{Simulation Results}
%\label{fig_sim}
%\end{figure}

% Note that IEEE typically puts floats only at the top, even when this
% results in a large percentage of a column being occupied by floats.

% An example of a double column floating figure using two subfigures.
% (The subfig.sty package must be loaded for this to work.)
% The subfigure \label commands are set within each subfloat command, the
% \label for the overall figure must come after \caption.
% \hfil must be used as a separator to get equal spacing.
% The subfigure.sty package works much the same way, except \subfigure is
% used instead of \subfloat.
%
%\begin{figure*}[!t]
%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
%\label{fig_first_case}}
%\hfil
%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
%\label{fig_second_case}}}
%\caption{Simulation results}
%\label{fig_sim}
%\end{figure*}
%
% Note that often IEEE papers with subfigures do not employ subfigure
% captions (using the optional argument to \subfloat), but instead will
% reference/describe all of them (a), (b), etc., within the main caption.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
%\begin{table}[!t]
%% increase table row spacing, adjust to taste
%\renewcommand{\arraystretch}{1.3}
% if using array.sty, it might be a good idea to tweak the value of
% \extrarowheight as needed to properly center the text within the cells
%\caption{An Example of a Table}
%\label{table_example}
%\centering
%% Some packages, such as MDW tools, offer better commands for making tables
%% than the plain LaTeX2e tabular which is used here.
%\begin{tabular}{|c||c|}
%\hline
%One & Two\\
%\hline
%Three & Four\\
%\hline
%\end{tabular}
%\end{table}

% Note that IEEE does not put floats in the very first column - or typically
% anywhere on the first page for that matter. Also, in-text middle ("here")
% positioning is not used. Most IEEE journals use top floats exclusively.
% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
% floats. This can be corrected via the \fnbelowfloat command of the
% stfloats package.

%\section{Fazit}
\section{Conclusion}
Put the conclusions of the work here. The conclusion is like the abstract with
an additional discussion of open points.

% if have a single appendix:
%\appendix[Proof of the Zonklar Equations]
% or
%\appendix  % for no appendix heading
% do not use \section anymore after \appendix, only \section*
% is possibly needed

% use appendices with more than one appendix
% then use \section to start each appendix
% you must declare a \section before using any
% \subsection or using \label (\appendices by itself
% starts a section numbered zero.)
%

% \appendices
% \section{Proof of the First Zonklar Equation}
% Appendix one text goes here.

% you can choose not to have a title for an appendix
% if you want by leaving the argument blank
% \section{}
% Appendix two text goes here.

% use section* for acknowledgement
%\section*{Acknowledgment}

%The authors would like to thank...

% Can use something like this to put references on a page
% by themselves when using endfloat and the captionsoff option.
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

% trigger a \newpage just before the given reference
% number - used to balance the columns on the last page
% adjust value as needed - may need to be readjusted if
% the document is modified later
%\IEEEtriggeratref{8}
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

% can use a bibliography generated by BibTeX as a .bbl file
% BibTeX documentation can be easily obtained at:
% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
% The IEEEtran BibTeX style support page is at:
% http://www.michaelshell.org/tex/ieeetran/bibtex/
%\bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
%\bibliography{IEEEabrv,../bib/paper}
%
% <OR> manually copy in the resultant .bbl file
% set second argument of \begin to the number of references
% (used to reserve space for the reference number labels box)
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}

% biography section
% 
% If you have an EPS/PDF photo (graphicx package needed) extra braces are
% needed around the contents of the optional argument to biography to prevent
% the LaTeX parser from getting confused when it sees the complicated
% \includegraphics command within an optional argument. (You could create
% your own custom macro containing the \includegraphics command to make things
% simpler here.)
%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
% or if you just want to reserve a space for a photo:

%\begin{IEEEbiography}{Michael Shell}
%Biography text here.
%\end{IEEEbiography}

% if you will not have a photo at all:
%\begin{IEEEbiographynophoto}{John Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

%\begin{IEEEbiographynophoto}{Jane Doe}
%Biography text here.
%\end{IEEEbiographynophoto}

% You can push biographies down or up by placing
% a \vfill before or after them. The appropriate
% use of \vfill depends on what kind of text is
% on the last page and whether or not the columns
% are being equalized.

%\vfill

% Can be used to pull up biographies so that the bottom of the last one
% is flush with the other column.
%\enlargethispage{-5in}

% that's all folks
\end{document}

### 2014-04-22 paper.tex
replaced title and author
removed comments

@@
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE! 
-%% User assumes all risk.
-%% In no event shall IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%
-%% File list of work: IEEEtran.cls, IEEEtran_HOWTO.pdf, bare_adv.tex,
-%%                    bare_conf.tex, bare_jrnl.tex, bare_jrnl_compsoc.tex
-%%*************************************************************************
-
-\documentclass[10pt,        % Don't change the font size!
-               a4paper,     % Don't change the paper size!
-               journal,     % Journal paper format
-%               draft       % Enable this parameter to get a draft version.
+\documentclass[10pt,
+               a4paper,
+               journal,
                ]{IEEEtran}
@@
 \usepackage[T1]{fontenc}
-
-% *** CITATION PACKAGES ***
 \usepackage{cite}
@@
 \begin{document}
-%
-% paper title
-% can use linebreaks \\ within to get better formatting as desired
-\title{Seminar Topic}
-%
-%
-% author names and IEEE memberships
-% note positions of commas and nonbreaking spaces ( ~ ) LaTeX will not break
-% a structure at a ~ so this keeps an author's name from being broken across
-% two lines.
-% use \thanks{} to gain access to the first footnote area
-% a separate \thanks must be used for each paragraph as LaTeX2e's \thanks
-% was not built to handle multiple paragraphs
-%
-
-\author{John~Smith}% <-this % stops a space
-%\thanks{M. Shell is with the Department
-%of Electrical and Computer Engineering, Georgia Institute of Technology, Atlanta,
-%GA, 30332 USA e-mail: (see http://www.michaelshell.org/contact.html).}% <-this % stops a space
-%\thanks{J. Doe and J. Doe are with Anonymous University.}% <-this % stops a space
-%\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}
-
-% note the % following the last \IEEEmembership and also \thanks - 
-% these prevent an unwanted space from occurring between the last author name
-% and the end of the author line. i.e., if you had this:
-% 
-% \author{....lastname \thanks{...} \thanks{...} }
-%                     ^------------^------------^----Do not want these spaces!
-%
-% a space would be appended to the last name and could cause every name on that
-% line to be shifted left slightly. This is one of those "LaTeX things". For
-% instance, "\textbf{A} \textbf{B}" will typeset as "A B" not "AB". To get
-% "AB" then you have to do: "\textbf{A}\textbf{B}"
-% \thanks is no different in this regard, so shield the last } of each \thanks
-% that ends a line with a % and do not let a space in before the next \thanks.
-% Spaces after \IEEEmembership other than the last one are OK (and needed) as
-% you are supposed to have spaces between the names. For what it is worth,
-% this is a minor point as most people would not even notice if the said evil
-% space somehow managed to creep in.
-
-% The paper headers
-%\markboth{Hauptseminar Sicherheit in der Informationstechnik, Sommersemester 2011}%
-\markboth{Advanced Seminar for Security in Information Technology, Summer Term 2011}%
-{John Smith: Seminar Topic}
-
-%\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
-%{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Journals}
-% The only time the second header will appear is for the odd numbered pages
-% after the title page when using the twoside option.
-% 
-% *** Note that you probably will NOT want to include the author's ***
-% *** name in the headers of peer review papers.                   ***
-% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
-% you desire.
-
-% If you want to put a publisher's ID mark on the page you can do it like
-% this:
-%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
-% Remember, if you use this you must call \IEEEpubidadjcol in the second
-% column for its text to clear the IEEEpubid mark.
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
+	\title{Constraint Programming - Inside Gecode}
+	\author{Benedikt~Schmidt}
+	\markboth{Advanced Seminar for Security in Information Technology, Summer Term 2014}%
+	{Benedikt Schmidt: Constraint Programming - Inside Gecode}	
 	\maketitle
@@
 	\begin{abstract}	
-%\boldmath
 		Put the abstract here. The abstract should be 200-300 words long and must
@@
 		topic in a broader context. 	
+	\end{abstract}
 
-\end{abstract}
-% IEEEtran.cls defaults to using nonbold math in the Abstract.
-% This preserves the distinction between vectors and scalars. However,
-% if the journal you are submitting to favors bold math in the abstract,
-% then you can use LaTeX's standard command \boldmath at the very start
-% of the abstract to achieve this. Many IEEE journals frown on math
-% in the abstract anyway.
-
-% Note that keywords are not normally used for peerreview papers.
-%\begin{IEEEkeywords}
-%IEEEtran, journal, \LaTeX, paper, template.
-%\end{IEEEkeywords}
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
-% \IEEEpeerreviewmaketitle
-
-%\section{Einleitung}
 	\section{Introduction}
@@
 
-%\section{Stand der Technik}
 	\section{State of the Art}
@@
 	Subsection text here.
-
-% needed in second column of first page if using \IEEEpubid
-%\IEEEpubidadjcol
 
@@
 
-% An example of a floating figure using the graphicx package.
-% Note that \label must occur AFTER (or within) \caption.
-% For figures, \caption should occur after the \includegraphics.
-% Note that IEEEtran v1.7 and later has special internal code that
-% is designed to preserve the operation of \label within \caption
-% even when the captionsoff option is in effect. However, because
-% of issues like this, it may be the safest practice to put all your
-% \label just after \caption rather than within \caption{}.
-%
-% Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
-% option should be used if it is desired that the figures are to be
-% displayed while in draft mode.
-%
-%\begin{figure}[!t]
-%\centering
-%\includegraphics[width=2.5in]{myfigure}
-% where an .eps filename suffix will be assumed under latex, 
-% and a .pdf suffix will be assumed for pdflatex; or what has been declared
-% via \DeclareGraphicsExtensions.
-%\caption{Simulation Results}
-%\label{fig_sim}
-%\end{figure}
-
-% Note that IEEE typically puts floats only at the top, even when this
-% results in a large percentage of a column being occupied by floats.
-
-% An example of a double column floating figure using two subfigures.
-% (The subfig.sty package must be loaded for this to work.)
-% The subfigure \label commands are set within each subfloat command, the
-% \label for the overall figure must come after \caption.
-% \hfil must be used as a separator to get equal spacing.
-% The subfigure.sty package works much the same way, except \subfigure is
-% used instead of \subfloat.
-%
-%\begin{figure*}[!t]
-%\centerline{\subfloat[Case I]\includegraphics[width=2.5in]{subfigcase1}%
-%\label{fig_first_case}}
-%\hfil
-%\subfloat[Case II]{\includegraphics[width=2.5in]{subfigcase2}%
-%\label{fig_second_case}}}
-%\caption{Simulation results}
-%\label{fig_sim}
-%\end{figure*}
-%
-% Note that often IEEE papers with subfigures do not employ subfigure
-% captions (using the optional argument to \subfloat), but instead will
-% reference/describe all of them (a), (b), etc., within the main caption.
-
-% An example of a floating table. Note that, for IEEE style tables, the 
-% \caption command should come BEFORE the table. Table text will default to
-% \footnotesize as IEEE normally uses this smaller font for tables.
-% The \label must come after \caption as always.
-%
-%\begin{table}[!t]
-%% increase table row spacing, adjust to taste
-%\renewcommand{\arraystretch}{1.3}
-% if using array.sty, it might be a good idea to tweak the value of
-% \extrarowheight as needed to properly center the text within the cells
-%\caption{An Example of a Table}
-%\label{table_example}
-%\centering
-%% Some packages, such as MDW tools, offer better commands for making tables
-%% than the plain LaTeX2e tabular which is used here.
-%\begin{tabular}{|c||c|}
-%\hline
-%One & Two\\
-%\hline
-%Three & Four\\
-%\hline
-%\end{tabular}
-%\end{table}
-
-% Note that IEEE does not put floats in the very first column - or typically
-% anywhere on the first page for that matter. Also, in-text middle ("here")
-% positioning is not used. Most IEEE journals use top floats exclusively.
-% Note that, LaTeX2e, unlike IEEE journals, places footnotes above bottom
-% floats. This can be corrected via the \fnbelowfloat command of the
-% stfloats package.
-
-%\section{Fazit}
 	\section{Conclusion}
@@
 
-% if have a single appendix:
-%\appendix[Proof of the Zonklar Equations]
-% or
-%\appendix  % for no appendix heading
-% do not use \section anymore after \appendix, only \section*
-% is possibly needed
-
-% use appendices with more than one appendix
-% then use \section to start each appendix
-% you must declare a \section before using any
-% \subsection or using \label (\appendices by itself
-% starts a section numbered zero.)
-%
-
-% \appendices
-% \section{Proof of the First Zonklar Equation}
-% Appendix one text goes here.
-
-% you can choose not to have a title for an appendix
-% if you want by leaving the argument blank
-% \section{}
-% Appendix two text goes here.
-
-% use section* for acknowledgement
-%\section*{Acknowledgment}
-
-%The authors would like to thank...
-
-% Can use something like this to put references on a page
-% by themselves when using endfloat and the captionsoff option.
-\ifCLASSOPTIONcaptionsoff
-  \newpage
-\fi
-
-% trigger a \newpage just before the given reference
-% number - used to balance the columns on the last page
-% adjust value as needed - may need to be readjusted if
-% the document is modified later
-%\IEEEtriggeratref{8}
-% The "triggered" command can be changed if desired:
-%\IEEEtriggercmd{\enlargethispage{-5in}}
-
-% references section
-
-% can use a bibliography generated by BibTeX as a .bbl file
-% BibTeX documentation can be easily obtained at:
-% http://www.ctan.org/tex-archive/biblio/bibtex/contrib/doc/
-% The IEEEtran BibTeX style support page is at:
-% http://www.michaelshell.org/tex/ieeetran/bibtex/
-%\bibliographystyle{IEEEtran}
-% argument is your BibTeX string definitions and bibliography database(s)
-%\bibliography{IEEEabrv,../bib/paper}
-%
-% <OR> manually copy in the resultant .bbl file
-% set second argument of \begin to the number of references
-% (used to reserve space for the reference number labels box)
 	\begin{thebibliography}{1}
-
 		\bibitem{IEEEhowto:kopka}
@@
 		  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
 	\end{thebibliography}
-
-% biography section
-% 
-% If you have an EPS/PDF photo (graphicx package needed) extra braces are
-% needed around the contents of the optional argument to biography to prevent
-% the LaTeX parser from getting confused when it sees the complicated
-% \includegraphics command within an optional argument. (You could create
-% your own custom macro containing the \includegraphics command to make things
-% simpler here.)
-%\begin{biography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{mshell}}]{Michael Shell}
-% or if you just want to reserve a space for a photo:
-
-%\begin{IEEEbiography}{Michael Shell}
-%Biography text here.
-%\end{IEEEbiography}
-
-% if you will not have a photo at all:
-%\begin{IEEEbiographynophoto}{John Doe}
-%Biography text here.
-%\end{IEEEbiographynophoto}
-
-% insert where needed to balance the two columns on the last page with
-% biographies
-%\newpage
-
-%\begin{IEEEbiographynophoto}{Jane Doe}
-%Biography text here.
-%\end{IEEEbiographynophoto}
-
-% You can push biographies down or up by placing
-% a \vfill before or after them. The appropriate
-% use of \vfill depends on what kind of text is
-% on the last page and whether or not the columns
-% are being equalized.
-
-%\vfill
-
-% Can be used to pull up biographies so that the bottom of the last one
-% is flush with the other column.
-%\enlargethispage{-5in}
-
-% that's all folks
 \end{document}

### 2014-04-22 paper.tex
removed unused section
defined basic structure

@@
 	\begin{abstract}	
-		Put the abstract here. The abstract should be 200-300 words long and must
-		answer the following questions:
-		\begin{itemize}
-		   \item Why is this topic important and interesting?
-		   \item What is the topic?
-		\end{itemize}
-		It's a good writing style to answer the "why" question first by putting the
-		topic in a broader context. 	
+		\textit{last thing to write}	
 	\end{abstract}
@@
 	\section{Introduction}
-	In general a scientific paper starts with an introduction to the topic.
+	\textit{something general on constraint programming}
 
 	\section{State of the Art}
-	In this section an analysis of the available literature on the topic is done.
-	This section may be split or subdivided into several sections or subsections.
+	\textit{what is in gecode implemented}
 
-	\subsection{Subsection Heading Here}
-	Subsection text here.
-
-	\subsubsection{Subsubsection Heading Here}
-	Subsubsection text here.
-
-	\section{Please Note}
-	Regular research papers need at least two additional sections here. One section
-	for contributions and methods and one section for the results. For seminar
-	papers these sections can be omitted.
+	\textit{how does the stuff in gecode work}
 
 	\section{Conclusion}
-	Put the conclusions of the work here. The conclusion is like the abstract with
-	an additional discussion of open points.
+	\textit{conclusion}
 

### 2014-05-02 paper.tex
introduction

@@
 	\section{Introduction}
-	\textit{something general on constraint programming}
+	Constraint programming is a tool to solve mathematical problems through the formulation of constraints, which describe the desired solution.
+As typically a foraml and exact definition, like it can be found in \cite[p.
+16]{handbookCP} for constraint propagation, is not as helpful as an example, I would like to give you a short overview on the topic through the well known problem \emph{SEND MORE MONEY}.
+
+	Consider the problem
+	\[SEND + MORE = MONEY\]
+	where every character is a variable and the position in the word describes the weight.
+	\begin{equation}
+	\begin{split}
+		S \cdot 10^3 + E \cdot 10^2 + N \cdot 10^1 + D \cdot 10^0 + \\ 
+		M \cdot 10^3 + O \cdot 10^2 + R \cdot 10^1 + E \cdot 10^0 = \\ 
+		M \cdot 10^4 + O \cdot 10^3 + N \cdot 10^2 + E \cdot 10^1 + Y \cdot 10^0
+	\end{split}
+	\end{equation}
+
+	To narrow the solution space a little bit down I state additional constraints:
+	\begin{enumerate}
+	\item All variables must have different values.
+		\begin{equation}
+			S \ne E \land S \ne N \land S \ne D \land \dots \land E \ne N \land E \ne D \dots
+		\end{equation}	
+	\item There must not be a trailing zero. Combined with the so-called all-different constraint \cite{allDifferent} from above this results in
+		\begin{equation}
+			S \ne 0 \land M \ne 0
+		\end{equation}
+	\end{enumerate}
+
+	Now a solver for constraint programming like Gecode provides the necessary tools to define these constraints in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected search algorithm.
+Actually this example contains only linear constraints and could therefore be solved through linear programming.
+As linear programming is only a subset of constraint programming the latter one is often the more useful one, because not every problem in the real world is linear or can be linearized in a useful way.
+Consequently constraint programming with its ability to consider constraints in there most general form can be used in various fields like scheduling of resources, computer-aided design and robotics \cite[p.
+221]{trendsInCP}.
+
+	In the following I will describe the necessary steps to solve a problem with constraint programming and as example of an implementation I will always refer to \emph{Gecode} \cite{gecode}.
 
 	\section{State of the Art}
-	\textit{what is in gecode implemented}
+	\subsection{Variable Domains}
 
-	\textit{how does the stuff in gecode work}
+	\subsection{Constraint Types}
+
+	\subsection{Constraint Propagation}
+
+	\subsection{Search Algorithms}
 
@@
 	\begin{thebibliography}{1}
-		\bibitem{IEEEhowto:kopka}
-		H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-		  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
+		\bibitem{handbookCP}
+		F.~Rossi, P.van~Beek and T.Walsh, \emph{Handbook of Constraint Programming}, Amsterdam, The Netherlands: Elsevier, 2006, ISBN: 978-0-444-52726-4
+		\bibitem{allDifferent}
+		W.~van~Hoeve, \emph{The Alldifferent Constraint: A Survey}, http://www.andrew.cmu.edu/user/vanhoeve/papers/alldiff.pdf
+		\bibitem{trendsInCP}
+		F.~Benhamou, N.~Jussien, B.A.O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN: 978-1-905209-97-2
+		\bibitem{gecode}
+		http://www.gecode.org/
 	\end{thebibliography}

### 2014-05-02 paper.tex
variable domains

@@
 \usepackage{cite}
+\usepackage{amsfonts}
 \usepackage[pdftex]{graphicx}
@@
 	\subsection{Variable Domains}
+	In constraint programming to a variable a set of possible values is connected, the so called domain.
+This domain can be a finite or inifite set of integers, true or false and intervals.
+Depending on the domain a variable is called an integer, boolean or floating variable.
+Possible ranges for a variable $x$ of each domain are:
+	\begin{itemize}
+		\item[Integer]: $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$, \dots
+		\item[Bool]: $x \in \{\text{true}, \text{false}\}$
+		\item[Float]: $x \in [-4, 3]$, $x \in [-10, 8] \cup [10, 12]$, \dots
+	\end{itemize}
+	This distinction is important as especially for floating variables the classical approach to constraint programming, a backtrack search, can not be applied directly.
+For integer and boolean variables the domain for each variable is reduced step by step during a search, either by constraints which rule out certain values or by trial and error.
+As floating variables do not have discrete values as their domains the procedere in this case is to reduce or split intervals based on constraints or to apply trial and error through splitting the intervals.
+I will come back to this topic later during the discussion of the branch-and-bound search.
 

### 2014-05-02 paper.tex
variable domains

@@
 For integer and boolean variables the domain for each variable is reduced step by step during a search, either by constraints which rule out certain values or by trial and error.
-As floating variables do not have discrete values as their domains the procedere in this case is to reduce or split intervals based on constraints or to apply trial and error through splitting the intervals.
+As floating variables do not have discrete values for their domains the procedere in this case is to reduce or split intervals based on constraints or to apply trial and error through splitting the intervals.
 I will come back to this topic later during the discussion of the branch-and-bound search.

### 2014-05-02 paper.tex
fixed itemize
added citation

--- 2014-05-02 paper.tex
@@
 	\begin{itemize}
-		\item[Integer]: $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$, \dots
-		\item[Bool]: $x \in \{\text{true}, \text{false}\}$
-		\item[Float]: $x \in [-4, 3]$, $x \in [-10, 8] \cup [10, 12]$, \dots
+		\item Integer: $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$, \dots
+		\item Bool: $x \in \{\text{true}, \text{false}\}$
+		\item Float: $x \in [-4, 3]$, $x \in [-10, 8] \cup [10, 12]$, \dots
 	\end{itemize}
@@
 For integer and boolean variables the domain for each variable is reduced step by step during a search, either by constraints which rule out certain values or by trial and error.
-As floating variables do not have discrete values as their domains the procedere in this case is to reduce or split intervals based on constraints or to apply trial and error through splitting the intervals.
+As floating variables do not have discrete values for their domains the procedere in this case is to reduce or split intervals based on constraints or to apply trial and error through splitting the intervals \cite[p.
+571]{handbookCP}.
 I will come back to this topic later during the discussion of the branch-and-bound search.

### 2014-05-02 paper.tex
constraint types

@@
 	\begin{equation}
+	\label{eq:linearConstraint}
 	\begin{split}
@@
 	\subsection{Constraint Types}
+	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \ref{eq:linearConstraint}.
+Which statements are possible to make varies than very much from the programming language in which the problem is implemented and the used toolset.
+Typical constraints would be
+	\begin{itemize}
+		\item Equalities: $x = y + 2$, $x^2 + y^2 = 1$, \dots
+		\item Inequalities: $x \ne y$, \dots
+		\item Relations: $x < 2 y$, $e^x < 4$, \dots
+	\end{itemize}
 

### 2014-05-02 paper.tex
added example of non-overlapping rectangles

@@
 \hyphenation{op-tical net-works semi-conduc-tor}
+\usepackage{listings}
+\usepackage{color}
+
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+
+\lstset{frame=tb,
+  language=C++,
+  aboveskip=3mm,
+  belowskip=3mm,
+  showstringspaces=false,
+  columns=flexible,
+  basicstyle={\small\ttfamily},
+  numbers=none,
+  numberstyle=\tiny\color{gray},
+  keywordstyle=\color{blue},
+  commentstyle=\color{dkgreen},
+  stringstyle=\color{mauve},
+  breaklines=true,
+  breakatwhitespace=true
+  tabsize=3
+}
 
@@
 	\begin{split}
-		S \cdot 10^3 + E \cdot 10^2 + N \cdot 10^1 + D \cdot 10^0 + \\ 
-		M \cdot 10^3 + O \cdot 10^2 + R \cdot 10^1 + E \cdot 10^0 = \\ 
-		M \cdot 10^4 + O \cdot 10^3 + N \cdot 10^2 + E \cdot 10^1 + Y \cdot 10^0
+		& S \cdot 10^3 + E \cdot 10^2 + N \cdot 10^1 + D \cdot 10^0 + \\ 
+		& M \cdot 10^3 + O \cdot 10^2 + R \cdot 10^1 + E \cdot 10^0 = \\ 
+		& M \cdot 10^4 + O \cdot 10^3 + N \cdot 10^2 + E \cdot 10^1 + Y \cdot 10^0
 	\end{split}
@@
 		\begin{equation}
-			S \ne E \land S \ne N \land S \ne D \land \dots \land E \ne N \land E \ne D \dots
+		\begin{split}
+			S \ne E \land S \ne N \land & S \ne D \land \dots \land \\
+			E \ne N \land & E \ne D \land \dots \land \\
+			& N \ne D \land \dots
+		\end{split}
 		\end{equation}	
@@
 	\end{itemize}
+	In Gecode for example it is very easy to create for example a nonlinear constraint based on the integer variables a, b, c and d \cite[p. 120]{programmingGecode}:
+	\begin{lstlisting}
+rel(home, a+b*(c+d) == 0);
+	\end{lstlisting}
+	This single statement in C++ implements the constraint
+	\begin{equation}
+		a + b \cdot (c + d) == 0
+	\end{equation}
+	what is a quite handy way to do it.
+
+	Hence there are boolean variables available it is also possible to use logical operators to combine other constraint types to logical expressions.
+To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap.
+This problem can be expressed mathematically by \cite[p.
+101]{programmingGecode}
+	\begin{equation}
+	\begin{split}
+		x_1 + d_1 \le x_2 & \lor x_2 + d_2 \le x_1 \lor \\
+		y_1 + d_1 \le y_2 & \lor y_2 + d_2 \le y_1
+	\end{split}
+	\end{equation}
+	and in Gecode by
+	\begin{lstlisting}
+rel(home,	(x1 + d1 <= x2) || 
+		(x2 + d2 <= x1) || 
+		(y1 + d1 <= y2) || 
+		(y2 + d2 <= y1));
+	\end{lstlisting}
+	which is again quite obvious to a reader with some experience in a C-like programming language.
 
@@
 		http://www.gecode.org/
+		\bibitem{programmingGecode}
+		C.~Schulte, G.~Tack and M.Z.Lagerkvist, \emph{Modeling and Programming with Gecode}, http://www.gecode.org/doc-latest/MPG.pdf
 	\end{thebibliography}

### 2014-05-02 paper.tex
added figure for squares

@@
 \hyphenation{op-tical net-works semi-conduc-tor}
+\usepackage{tikz}
 \usepackage{listings}
@@
 	\subsection{Constraint Types}
-	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \ref{eq:linearConstraint}.
+	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like [equation \ref{eq:linearConstraint}].
 Which statements are possible to make varies than very much from the programming language in which the problem is implemented and the used toolset.
@@
 
+	\begin{figure}
+	\center
+	\begin{tikzpicture}
+		\draw (0, 0) rectangle (1, 1);
+		\node at (0, -0.3) {($x_1$, $y_1$)};
+		\draw[arrows=<->] (1.2, 0) -- (1.2, 1);
+		\node at (1.5, 0.5) {$d_1$};
+		\draw (1.5, 1.5) rectangle(2, 2);
+		\node at (1.5, 1.2) {($x_2$, $y_2$)};
+		\draw[arrows=<->] (2.2, 1.5) -- (2.2, 2);
+		\node at (2.5, 1.75) {$d_2$};
+	\end{tikzpicture}
+	\caption{two squares which should not overlap}
+	\label{fig:squares}
+	\end{figure}
 	Hence there are boolean variables available it is also possible to use logical operators to combine other constraint types to logical expressions.
-To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap.
+To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap [figure \ref{fig:squares}].
 This problem can be expressed mathematically by \cite[p.

### 2014-05-03 paper.tex
correction
constraint propagation

@@
 	\end{equation}
-	what is a quite handy way to do it.
+	which is a quite handy way to do it.
 
@@
 	\subsection{Constraint Propagation}
+	Constraint propagation is a key concept in constraint programming as it allows to speed-up the search significantly.
+The term propagation refers in this area to the reduction of domains of variables which are infeasible, considering other constraints and affected variable domains.
+As the size of the variable domain is related to possibilities which have to be checked by a search, the constraint propagation is able to prune infeasible subtrees.
+Therefore the whole search process is accelerated.
+
+	The actual implementation of constraint propagation differs again from one toolset to another, but the basic concepts are for sure the same.
+As the topic of this paper is the implementation in Gecode I will concentrate on this one.
+
+	Constraints in Gecode are always evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the subtrees.
+The problem here is to decide in which order the constraints should be considered, and it may even be necessary to evaluate some constraints more often.
+A short example will explain this very well \cite[p.
+575]{handbookCP}: Let there be two constraints
+	\begin{equation}
+	c_1: y - x^2 = 0
+	\end{equation}
+	\begin{equation}
+	c_2: y - x - 1 = 0
+	\end{equation}
+	and let the variable domains be $x, y \in [-4, 4]$. Starting with the evaluation of constraint one follows a chain of reductions of the variable domains, based only on constraint propagation:
+	\begin{equation}
+	\begin{split}
+	c_1:\ &y \in [-4, 4] \rightarrow y \in [0, 4] \\
+	c_1:\ &x \in [-4, 4] \rightarrow x \in [-2, 2] \\
+	c_2:\ &y \in [0, 4] \rightarrow y \in [0, 3] \\
+	c_2:\ &x \in [-2, 2] \rightarrow x \in [-1, 2] \\
+	c_1:\ &x \in [-1, 2] \rightarrow x \in [-1, 1.73 \dots] \\
+	&\dots
+	\end{split}
+	\end{equation}
+	Actually this behaviour is even very nice to have as it narrows the down the search space very efficient.
+
+	To implement this repeated evaluation of constraints in Gecode all constraints subscribe to the domains of the variables which are part of them.
+With this method the constraints get noticed whenever a variable domain is reduced and they can signal the system that they need to be evaluated once again.
+
+	The next difficult decision is which constraint should evaluated first.
+In the example above the constraints were quite simple and therefore it won't make a difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time.
+Unfortunately not all constraints are that easy to evaluate and for this case Gecode sets for every constraint a certain cost \cite[p.
+275]{programmingGecode}.
+Based on this cost the constraints are sorted into buckets which are then iterated from the constraints with lower costs to the ones with higher costs.
+These means, whenever a variable domain is changed the first try is to use constraints with the lowest possible cost and only if none of them needs to be evaluated constraints with higher costs are considered.
 

### 2014-05-03 paper.tex
constraint propagation search tree

@@
 \hyphenation{op-tical net-works semi-conduc-tor}
+\usepackage{stmaryrd}
+\usepackage{bbding}
 \usepackage{tikz}
+\usetikzlibrary{shapes}
+\usetikzlibrary{calc}
+\usetikzlibrary{positioning}
 \usepackage{listings}
@@
 
+\newcommand{\reffig}[1]{{[Fig.~\ref{#1}.]}}
+\newcommand{\refeq}[1]{{[Eq.~\ref{#1}.]}}
+
 \begin{document}
@@
 	Constraint programming is a tool to solve mathematical problems through the formulation of constraints, which describe the desired solution.
-As typically a foraml and exact definition, like it can be found in \cite[p.
-16]{handbookCP} for constraint propagation, is not as helpful as an example, I would like to give you a short overview on the topic through the well known problem \emph{SEND MORE MONEY}.
+As typically a foraml and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint propagation, is not as helpful as an example, I would like to give you a short overview on the topic through the well known problem \emph{SEND MORE MONEY}.
 
@@
 As linear programming is only a subset of constraint programming the latter one is often the more useful one, because not every problem in the real world is linear or can be linearized in a useful way.
-Consequently constraint programming with its ability to consider constraints in there most general form can be used in various fields like scheduling of resources, computer-aided design and robotics \cite[p.
-221]{trendsInCP}.
+Consequently constraint programming with its ability to consider constraints in there most general form can be used in various fields like scheduling of resources, computer-aided design and robotics \cite[p.~221]{trendsInCP}.
 
@@
 For integer and boolean variables the domain for each variable is reduced step by step during a search, either by constraints which rule out certain values or by trial and error.
-As floating variables do not have discrete values for their domains the procedere in this case is to reduce or split intervals based on constraints or to apply trial and error through splitting the intervals \cite[p.
-571]{handbookCP}.
+As floating variables do not have discrete values for their domains the procedere in this case is to reduce or split intervals based on constraints or to apply trial and error through splitting the intervals \cite[p.~571]{handbookCP}.
 I will come back to this topic later during the discussion of the branch-and-bound search.
@@
 	\subsection{Constraint Types}
-	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like [equation \ref{eq:linearConstraint}].
+	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \refeq{eq:linearConstraint}.
 Which statements are possible to make varies than very much from the programming language in which the problem is implemented and the used toolset.
@@
 	\end{itemize}
-	In Gecode for example it is very easy to create for example a nonlinear constraint based on the integer variables a, b, c and d \cite[p. 120]{programmingGecode}:
+	In Gecode for example it is very easy to create for example a nonlinear constraint based on the integer variables a, b, c and d \cite[p.~120]{programmingGecode}:
 	\begin{lstlisting}
@@
 	Hence there are boolean variables available it is also possible to use logical operators to combine other constraint types to logical expressions.
-To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap [figure \ref{fig:squares}].
+To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap \reffig{fig:squares}.
 This problem can be expressed mathematically by \cite[p.
@@
 
+	\begin{figure}
+	\center
+	\begin{tikzpicture}
+		\tikzstyle{arrow}=[draw, -latex]
+		\node[rectangle,draw] at (0, 3) (levelOne) {$x \in \{1, 5, 6\} \land y \in \{1, 2\}$};
+
+		\node[rectangle,draw] at (-3, 2) (levelTwoElementOne) {$x = 1$};
+		\coordinate[left=1.5cm of levelOne.south] (levelOnePointOne);
+  		\path[arrow] (levelTwoElementOne) -- (levelOnePointOne);
+
+		\node[rectangle,draw] at (0, 2) (levelTwoElementTwo) {$x = 5$};
+		\coordinate[left=0cm of levelOne.south] (levelOnePointTwo);
+  		\path[arrow] (levelTwoElementTwo) -- (levelOnePointTwo);
+
+		\node[rectangle,draw] at (3, 2) (levelTwoElementThree) {$x = 6$};
+		\coordinate[right=1.5cm of levelOne.south] (levelOnePointThree);
+  		\path[arrow] (levelTwoElementThree) -- (levelOnePointThree);
+
+		\node[rectangle,draw] at (-3.6, 1) (levelThreeElementOne) {$y = 1$};
+		\coordinate[left=0.2cm of levelTwoElementOne.south] (levelTwoElementOnePointOne);
+  		\path[arrow] (levelThreeElementOne) -- (levelTwoElementOnePointOne);
+		\node[rectangle,draw] at (-3.6, 0) (levelFourElementOne) {$c_2 \lightning$};
+  		\path[arrow] (levelFourElementOne) -- (levelThreeElementOne);
+		\node[rectangle,draw] at (-2.4, 1) (levelThreeElementTwo) {$y = 2$};
+		\coordinate[right=0.2cm of levelTwoElementOne.south] (levelTwoElementOnePointTwo);
+  		\path[arrow] (levelThreeElementTwo) -- (levelTwoElementOnePointTwo);
+		\node[rectangle,draw] at (-2.4, 0) (levelFourElementTwo) {\Checkmark};
+  		\path[arrow] (levelFourElementTwo) -- (levelThreeElementTwo);
+
+		\node[rectangle,draw] at (-0.6, 1) (levelThreeElementThree) {$y = 1$};
+		\coordinate[left=0.2cm of levelTwoElementTwo.south] (levelTwoElementTwoPointOne);
+  		\path[arrow] (levelThreeElementThree) -- (levelTwoElementTwoPointOne);
+		\node[rectangle,draw] at (-0.6, 0) (levelFourElementThree) {$c_1 \lightning$};
+  		\path[arrow] (levelFourElementThree) -- (levelThreeElementThree);
+		\node[rectangle,draw] at (0.6, 1) (levelThreeElementFour) {$y = 2$};
+		\coordinate[right=0.2cm of levelTwoElementTwo.south] (levelTwoElementTwoPointTwo);
+  		\path[arrow] (levelThreeElementFour) -- (levelTwoElementTwoPointTwo);
+		\node[rectangle,draw] at (0.6, 0) (levelFourElementFour) {$c_1 \lightning$};
+  		\path[arrow] (levelFourElementFour) -- (levelThreeElementFour);
+
+		\node[rectangle,draw] at (2.4, 1) (levelThreeElementFive) {$y = 1$};
+		\coordinate[left=0.2cm of levelTwoElementThree.south] (levelTwoElementThreePointOne);
+  		\path[arrow] (levelThreeElementFive) -- (levelTwoElementThreePointOne);
+		\node[rectangle,draw] at (2.4, 0) (levelFourElementFive) {$c_1 \lightning$};
+  		\path[arrow] (levelFourElementFive) -- (levelThreeElementFive);
+		\node[rectangle,draw] at (3.6, 1) (levelThreeElementSix) {$y = 2$};
+		\coordinate[right=0.2cm of levelTwoElementThree.south] (levelTwoElementThreePointTwo);
+  		\path[arrow] (levelThreeElementSix) -- (levelTwoElementThreePointTwo);
+		\node[rectangle,draw] at (3.6, 0) (levelFourElementSix) {$c_1 \lightning$};
+  		\path[arrow] (levelFourElementSix) -- (levelThreeElementSix);
+	\end{tikzpicture}
+	\caption{search tree without constraint propagation}
+	\label{fig:bigTree}
+	\end{figure}
+
+	To visualize this effect consider a constraint problem based on two variables $x \in \{1, 5, 6\}$ and $y \in \{1, 2\}$ and two constraints based on them.
+	\begin{equation}
+		c_1:\ x \le y
+	\end{equation}
+	\begin{equation}
+		c_2:\ x + y = 3
+	\end{equation}
+	Without constraint propagation and branching first on $x$ this results in search tree \reffig{fig:bigTree}.
+
 	The actual implementation of constraint propagation differs again from one toolset to another, but the basic concepts are for sure the same.
@@
 The problem here is to decide in which order the constraints should be considered, and it may even be necessary to evaluate some constraints more often.
-A short example will explain this very well \cite[p.
-575]{handbookCP}: Let there be two constraints
+A short example will explain this very well \cite[p.~575]{handbookCP}: Let there be two constraints
 	\begin{equation}
@@
 In the example above the constraints were quite simple and therefore it won't make a difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time.
-Unfortunately not all constraints are that easy to evaluate and for this case Gecode sets for every constraint a certain cost \cite[p.
-275]{programmingGecode}.
+Unfortunately not all constraints are that easy to evaluate and for this case Gecode sets for every constraint a certain cost \cite[p.~275]{programmingGecode}.
 Based on this cost the constraints are sorted into buckets which are then iterated from the constraints with lower costs to the ones with higher costs.

### 2014-05-03 paper.tex
constraint propagation search tree

@@
   		\path[arrow] (levelThreeElementOne) -- (levelTwoElementOnePointOne);
-		\node[rectangle,draw] at (-3.6, 0) (levelFourElementOne) {$c_2 \lightning$};
+		\node[rectangle,draw] at (-3.6, 0) (levelFourElementOne) {\Checkmark};
   		\path[arrow] (levelFourElementOne) -- (levelThreeElementOne);
@@
 
-	To visualize this effect consider a constraint problem based on two variables $x \in \{1, 5, 6\}$ and $y \in \{1, 2\}$ and two constraints based on them.
+	\begin{figure}
+	\center
+	\begin{tikzpicture}
+		\tikzstyle{arrow}=[draw, -latex]
+		\node[rectangle,draw] at (0, 2) (levelOne) {$x \in \{1\} \land y \in \{1, 2\}$};
+
+		\node[rectangle,draw] at (-1, 1) (levelTwoElementOne) {$y = 1$};
+		\coordinate[left=0.2cm of levelOne.south] (levelOnePointOne);
+  		\path[arrow] (levelTwoElementOne) -- (levelOnePointOne);
+		\node[rectangle,draw] at (-1, 0) (levelThreeElementOne) {\Checkmark};
+  		\path[arrow] (levelThreeElementOne) -- (levelTwoElementOne);
+		\node[rectangle,draw] at (1, 1) (levelTwoElementTwo) {$y = 2$};
+		\coordinate[right=0.2cm of levelOne.south] (levelOnePointTwo);
+  		\path[arrow] (levelTwoElementTwo) -- (levelOnePointTwo);
+		\node[rectangle,draw] at (1, 0) (levelThreeElementTwo) {\Checkmark};
+  		\path[arrow] (levelThreeElementTwo) -- (levelTwoElementTwo);
+	\end{tikzpicture}
+	\caption{search tree with constraint propagation}
+	\label{fig:smallTree}
+	\end{figure}
+
+	To visualize this effect consider a constraint problem based on two variables $x \in \{1, 5, 6\}$ and $y \in \{1, 2\}$ and one constraint based on them.
 	\begin{equation}
@@
 	\end{equation}
-	\begin{equation}
-		c_2:\ x + y = 3
-	\end{equation}
 	Without constraint propagation and branching first on $x$ this results in search tree \reffig{fig:bigTree}.
+Obviously the right parts of these tree can already be pruned at the root with propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these subtrees \reffig{fig:smallTree}.
 

### 2014-05-03 paper.tex
constraint propagation

@@
 	\subsection{Constraint Types}
+
+	\begin{figure}
+	\center
+	\begin{tikzpicture}
+		\draw (0, 0) rectangle (1, 1);
+		\node at (0, -0.3) {($x_1$, $y_1$)};
+		\draw[arrows=<->] (1.2, 0) -- (1.2, 1);
+		\node at (1.5, 0.5) {$d_1$};
+		\draw (1.5, 1.5) rectangle(2, 2);
+		\node at (1.5, 1.2) {($x_2$, $y_2$)};
+		\draw[arrows=<->] (2.2, 1.5) -- (2.2, 2);
+		\node at (2.5, 1.75) {$d_2$};
+	\end{tikzpicture}
+	\caption{two squares which should not overlap}
+	\label{fig:squares}
+	\end{figure}
+
 	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \refeq{eq:linearConstraint}.
@@
 
-	\begin{figure}
-	\center
-	\begin{tikzpicture}
-		\draw (0, 0) rectangle (1, 1);
-		\node at (0, -0.3) {($x_1$, $y_1$)};
-		\draw[arrows=<->] (1.2, 0) -- (1.2, 1);
-		\node at (1.5, 0.5) {$d_1$};
-		\draw (1.5, 1.5) rectangle(2, 2);
-		\node at (1.5, 1.2) {($x_2$, $y_2$)};
-		\draw[arrows=<->] (2.2, 1.5) -- (2.2, 2);
-		\node at (2.5, 1.75) {$d_2$};
-	\end{tikzpicture}
-	\caption{two squares which should not overlap}
-	\label{fig:squares}
-	\end{figure}
 	Hence there are boolean variables available it is also possible to use logical operators to combine other constraint types to logical expressions.
@@
 	Without constraint propagation and branching first on $x$ this results in search tree \reffig{fig:bigTree}.
-Obviously the right parts of these tree can already be pruned at the root with propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these subtrees \reffig{fig:smallTree}.
+Obviously the right parts of this tree can already be pruned already at the root by propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these subtrees \reffig{fig:smallTree}.
 
@@
 	\end{equation}
-	Actually this behaviour is even very nice to have as it narrows the down the search space very efficient.
+	Actually this behaviour is very nice to have as it narrows the down the search space efficiently.
 
@@
 Based on this cost the constraints are sorted into buckets which are then iterated from the constraints with lower costs to the ones with higher costs.
-These means, whenever a variable domain is changed the first try is to use constraints with the lowest possible cost and only if none of them needs to be evaluated constraints with higher costs are considered.
+This means, whenever a variable domain is changed, the first try is to use constraints with the lowest possible cost and only if none of them needs to be evaluated constraints with higher costs are considered.
 

### 2014-05-03 paper.tex
corrections

@@
 	\end{equation}
-	Actually this behaviour is very nice to have as it narrows the down the search space efficiently.
+	Actually this behaviour is very nice to have as it narrows efficiently the search space down.
 
@@
 
-	The next difficult decision is which constraint should evaluated first.
-In the example above the constraints were quite simple and therefore it won't make a difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time.
-Unfortunately not all constraints are that easy to evaluate and for this case Gecode sets for every constraint a certain cost \cite[p.~275]{programmingGecode}.
+	The next difficult decision is which constraint should be evaluated first.
+In the example above the constraints were quite simple and therefore it won't make a big difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time.
+Unfortunately not all constraints are that easy to evaluate and for this case Gecode defines for every constraint a certain cost \cite[p.~275]{programmingGecode}.
 Based on this cost the constraints are sorted into buckets which are then iterated from the constraints with lower costs to the ones with higher costs.

### 2014-05-03 paper.tex
search algorithms

@@
 	\subsection{Search Algorithms}
+	In constraint programming search algorithm define the way the search tree is constructed, starting from the root node.
+To improve the performance of the search usually at every newly created node a constraint propagation is executed to narrow the search space down.
+Based on these two steps then either one, the best or no solution is found.
+What the actual result in the end is depends obviously on the search algorithm.
+
+	In Gecode it is possible to define its own search strategies, but the user is supported at this point with some standard engines which can be used to build custom ones:
+	\begin{itemize}
+	\item Depth-First Search
+	\item Limited Discrepancy Search
+	\item Branch-and-Bound Search
+	\item Depth-First Search Restart Optimization
+	\end{itemize}
+
+	In the following part I will discuss these searches and some additional, well known techniques, which can be applied during searches in constraint programming.
+
+	\subsubsection{Depth-First Search}
+	The Depth-First Search is a good point to start, as it was also historical the first developed one.
+The idea behind this search algorithm is that every possible combination of variables is examined.
+Obviously in its basic form it can not be applied to problems with inifinite domains; for such problems with such variable domains it is necessary to use for example a Branch-and-Bound Search.
+
+	The starting point is the root node, from where one the algorithm is called recursively.
+At every node mainly two steps are executed: First a constraint propagation and then a branching, which divides the search space into two disjoint sub spaces, on which the same algorithm is applied again.
+The search stops in a leave if either no branching is possible anymore and therefore a valid solution is found, or one or more variable domains are emptied, which indicates that in this leave no possible solution exists.
+
+	This search is a so-called complete algorithm as it can be proven that all possible solutions will be found \cite[p.~85]{handbookCP}.
+Because of this characteristic it is very useful to proove for example that for a certain problem no solution exists.
+The main drawback obviously is the execution time, as it does an exhaustive search, and maybe even the memory usage if during the search many solutions are found.
+
+	\subsubsection{Depth-First Search with Optimization}
+	If in the constraint programming is a certain objective is defined which should be either minimized or maximized a modified version of the Depth-First Search can be applied.
+The only difference is that it evalutes the objective function for every found solution concerning the constraints and stores during the search only the best one.
+To accelerate the search after one solution is already found it is possible to add an additional constraint which defines, that the new objective must be greater than the one related to the best solution till that point.
+Through this enhancement it is possible to prune even more sub trees and therefore to get the desired result faster.
+
+	This algorithm is again a complete one, therefore it will always find the best solution if one exists.
+This algorithm does not inherit the bad memory characteristics from the Depth-First Search as only one solution is stored, but the search is still exhaustive and can be very slow, depending on the size of the variable domains.
+
+	\subsubsection{Limited Discrepancy Search}
+
+	\subsubsection{Branch-and-Bound Search}
+
+	\subsubsection{Branching Strategies}
+
+	\subsubsection{Backtracking}
+
+	\subsubsection{Backjumping}
+
+	\subsubsection{Restart Strategies}
 
@@
 		\bibitem{handbookCP}
-		F.~Rossi, P.van~Beek and T.Walsh, \emph{Handbook of Constraint Programming}, Amsterdam, The Netherlands: Elsevier, 2006, ISBN: 978-0-444-52726-4
+		F.~Rossi, P.~van~Beek and T.~Walsh, \emph{Handbook of Constraint Programming}, Amsterdam, The Netherlands: Elsevier, 2006, ISBN: 978-0-444-52726-4
 		\bibitem{allDifferent}
@@
 		\bibitem{trendsInCP}
-		F.~Benhamou, N.~Jussien, B.A.O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN: 978-1-905209-97-2
+		F.~Benhamou, N.~Jussien, B.~A.~O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN: 978-1-905209-97-2
 		\bibitem{gecode}
@@
 		\bibitem{programmingGecode}
-		C.~Schulte, G.~Tack and M.Z.Lagerkvist, \emph{Modeling and Programming with Gecode}, http://www.gecode.org/doc-latest/MPG.pdf
+		C.~Schulte, G.~Tack and M.~Z.~Lagerkvist, \emph{Modeling and Programming with Gecode}, http://www.gecode.org/doc-latest/MPG.pdf
+		\bibitem{limitedDiscrepancy}
+		W.~D.~Harvey, M.~L.~Ginsberg, \emph{Limited Discrepancy Search}, Morgan Kaufmann, 1995
 	\end{thebibliography}

### 2014-05-03 paper.tex
search algorithms

@@
 	\subsubsection{Limited Discrepancy Search}
+	Limited Discrepancy Search is a variation of a more general Best-First Search and was first mentioned by Harvey and Ginsberg \cite{limitedDiscrepancy}.
+The basic concept is that during the iteration through a search tree it is better, in terms of execution time or, more formal, complexity, to fail closer to the root.
+If one constraint fails very close to the root only litte effort was put into this branch.
 

### 2014-05-03 paper.tex
corrections

@@
 The idea behind this search algorithm is that every possible combination of variables is examined.
-Obviously in its basic form it can not be applied to problems with inifinite domains; for such problems with such variable domains it is necessary to use for example a Branch-and-Bound Search.
+Obviously in its basic form it can not be applied to problems with inifinite domains; for problems with such variable domains it is necessary to use for example a Branch-and-Bound Search.
 
@@
 
-	\subsubsection{Backtracking}
-
 	\subsubsection{Backjumping}

### 2014-05-03 paper.tex
search algorithms

--- 2014-05-03 paper.tex
@@
 The main drawback obviously is the execution time, as it does an exhaustive search, and maybe even the memory usage if during the search many solutions are found.
+
+	By stopping the search after one feasible solution is found the algorithm the search can often be accelerated.
+Although this slightly modified version is not complete anymore, in practice it is in some areas sufficient to know just one valid solution.
 
@@
 
-	\subsubsection{Limited Discrepancy Search}
-
 	\subsubsection{Branch-and-Bound Search}
@@
 
-	\subsubsection{Backtracking}
+	\subsubsection{Limited Discrepancy Search}
+	Limited Discrepancy Search falls into the category of Best-First searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy} and it depends on some heurisitics for the tree, which is determined through branching strategies.
+The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive and second heurisitics work worse the farther away from the solution they are.
+Consequently, in a chronological backtrack search the worst decisions are the most expensive ones.
+The Limited Discrepancy Search solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later.
+The improvement ahead of a chronological backtrack were shown by experimental results and theoretical proven in \cite{limitedDiscrepancy}.
 

### 2014-05-03 paper.tex
added pseudo code for LDS
added pseudo code for depth-first search

@@
 \interdisplaylinepenalty=2500
-\usepackage{algorithmic}
 \usepackage{array}
@@
 \usepackage{bbding}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
 \usepackage{tikz}
@@
 	\subsubsection{Depth-First Search}
-	The Depth-First Search is a good point to start, as it was also historical the first developed one.
+	\begin{algorithm}
+		\begin{algorithmic}
+			\Function{DFS}{node}
+				\If{$Goal(node)$}
+					\State \Return $node$
+				\EndIf
+				\State $s \gets Successors(node)$
+				\If{$Null(s)$}
+					\State \Return $nil$
+				\EndIf
+				\State $result \gets DFS(Left(s))$
+				\If{$result \ne nil$}
+					\State \Return $result$
+				\Else
+					\State \Return $DFS(Right(s))$
+				\EndIf
+			\EndFunction
+		\end{algorithmic}
+		\caption{Depth-first search in a binary search tree for one feasible solution, without constraint propagation}
+	\end{algorithm}
+
+	The Depth-first search is a good point to start, as it was also historical the first developed one.
 The idea behind this search algorithm is that every possible combination of variables is examined.
 Obviously in its basic form it can not be applied to problems with inifinite domains; for problems with such variable domains it is necessary to use for example a Branch-and-Bound Search.
+Sometimes the Depth-First Search is also referred to as chronological backtrack search.
 
@@
 
-	By stopping the search after one feasible solution is found the algorithm the search can often be accelerated.
+	By stopping the search after one feasible solution is found the algorithm the Depth-first search can often be accelerated.
 Although this slightly modified version is not complete anymore, in practice it is in some areas sufficient to know just one valid solution.
@@
 	\subsubsection{Depth-First Search with Optimization}
-	If in the constraint programming is a certain objective is defined which should be either minimized or maximized a modified version of the Depth-First Search can be applied.
+	If in the constraint programming is a certain objective is defined which should be either minimized or maximized a modified version of the Depth-first search can be applied.
 The only difference is that it evalutes the objective function for every found solution concerning the constraints and stores during the search only the best one.
@@
 	\subsubsection{Limited Discrepancy Search}
-	Limited Discrepancy Search falls into the category of Best-First searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy} and it depends on some heurisitics for the tree, which is determined through branching strategies.
+	Limited Discrepancy Search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy} and it depends on some heurisitics for the tree, which is determined through branching strategies.
 The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive and second heurisitics work worse the farther away from the solution they are.
@@
 The improvement ahead of a chronological backtrack were shown by experimental results and theoretical proven in \cite{limitedDiscrepancy}.
+
+	\begin{algorithm}
+		\begin{algorithmic}
+			\Function{$LDSProbe$}{$node$, $k$}
+				\If{$Goal(node)$}
+				    \State \Return $node$
+				\EndIf
+				\State $s \gets Successors(node)$
+				\If{$Null(s)$}
+					\State \Return $nil$
+				\EndIf
+				\If{$k = 0$}
+					\State \Return $LDSProbe(Left(s), 0)$
+				\Else
+					\State $result \gets LDSProbe(Right(s), k - 1)$
+					\If{$result \ne nil$}
+						\State \Return $result$
+					\EndIf
+					\State \Return $LDSProbe(Left(s), k)$
+				\EndIf
+			\EndFunction
+		\end{algorithmic}
+		\caption{Probe of Limited Discrepancy Search for a binary search tree from \cite{limitedDiscrepancy}, without constraint propagation}
+	\end{algorithm}
+
+	\begin{algorithm}
+		\begin{algorithmic}			
+			\Function{$LDS$}{$node$}
+				\For{$x \gets 0\ \text{to maximum depth}$}
+					\State $result \gets LDSProbe(node, x)$
+					\If{$result \ne nil$}
+						\State \Return $result$
+					\EndIf
+				\EndFor
+			\EndFunction
+		\end{algorithmic}
+		\caption{Limited Discrepancy Search for a binary search tree from \cite{limitedDiscrepancy}, without constraint propagation}
+	\end{algorithm}
 

### 2014-05-03 paper.tex
branching strategies

@@
 
+	\subsubsection{Branching Strategies}
+	If at one node it is not possible to empty one variable domain or to reduce all variable domains to the size one it is necessary to make a decision: How should be the next subtree be build?
+The question is answered by the selected branching strategy, which depends on the variable domain.
+For Integer (and Bool) Variables there are three possible strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the subtrees.
+All the strategies have in common that they create additional temporary constraints, which are only valid for the specific subtree.
+The temporay addition of constraints is called \emph{posting}.
+	\begin{itemize}
+	\item Enumeration: The domain for $x \in D = {x_1, x_2, \dots}$ is split up into $|D|$-branches where for every branch $i$ one additional constraint $c:\ x = x_i$ is posted.
+	\item Binary choice points: One value $x_i$ of the domain $D$ of the variable $x$ is selected and the domain is split up into two subtrees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted.
+This results into something like a binary tree at this point.
+	\item Domain splitting: The domain $D$ for the variable $x$ is split up into two disjoint intervals or sets, for example $x < 2$ and $x \ge 2$.
+	\end{itemize}
+
+	Obviously for variables with a Float domain the Enumeration and Binary choice points are not feasible, but the last one, domain splitting, is typically used.
+As termination criteria for a Float it is often necessary to define a epsilon which describes the interval length, at which the solver states to have found a solution.
+
+	For all branching strategies there is a big choice of freedom left, which can be used to implement some heuristics.
+This additional information is either gained during the search itself of pre defined by the user.
+
 	\subsubsection{Depth-First Search}
@@
 
-	\subsubsection{Branching Strategies}
-
 	\subsubsection{Limited Discrepancy Search}
-	Limited Discrepancy Search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy} and it depends on some heurisitics for the tree, which is determined through branching strategies.
+	Limited Discrepancy Search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy} and it depends on some heurisitics for the tree, which are determined through branching strategies.
 The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive and second heurisitics work worse the farther away from the solution they are.
@@
 The Limited Discrepancy Search solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later.
-The improvement ahead of a chronological backtrack were shown by experimental results and theoretical proven in \cite{limitedDiscrepancy}.
+The improvement compared to a chronological backtrack was shown by experimental results and theoretical proven in \cite{limitedDiscrepancy}.
 

### 2014-05-03 paper.tex
branch-and-bound

@@
 	\subsubsection{Branch-and-Bound Search}
+	Branch-and-bound refers actually to two parts, a Branch-and-reduce search and bounding.
+Branch-and-reduce itself is the continous constraint propagation to reduce the variable domains combined with branching when there is no further reduction by constraint propagation possible.
+A simple heuristic for branching would be to split the biggest domain into two equal sized intervals.
+	Bounding is then the step which transforms the Branch-and-reduce search into an optimization. To be able to do this an objective function is necessary, which can be inverted.
 

### 2014-05-03 paper.tex
backjumping

@@
 	\subsubsection{Backjumping}
+	A specific version of backjumping was already mentioned, the Limited Discrepancy Search, where instead of jumping one node back the algorithm always jumps as far back to the root node as possible.
+In general this is proven to be better than the chronological backtracking \cite{limitedDiscrepancy}, but for certain problems it may be possible to find better points to jump back.
+A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the deadend.
+How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is down wrong the backjumping can even be less efficient than chronological backtracking \cite[p.~100]{handbookCP}.
 

### 2014-05-04 paper.tex
branch and bound

@@
 A simple heuristic for branching would be to split the biggest domain into two equal sized intervals.
-	Bounding is then the step which transforms the Branch-and-reduce search into an optimization. To be able to do this an objective function is necessary, which can be inverted.
+	Bounding is then the step which transforms the Branch-and-reduce search into an optimization.
+To be able to do this an objective function is necessary.
+In the following I will assume that the objective function has to be minimized with a loss of generality.
+	Every Branch splits the total problem into two sub problems.
+With some additional constraints, for example one variable must be an integer $x \in \mathbb{Z}$, it may happen that the real optimum of a sub problem is not the same one as if this additional constraint is considered.
+For finding the first optimum the trick is to branch in ways that a certain sub problem ends up with only a few possible candidates left, which then can be tested each.
+With this first optimum the search algorithm has then a so far best optimum which can be used as upper bound for the optimum.
+If for one sub problem the relaxed optimum, the optimum without for example the constraint of integrity, is worse than this upper bound the sub problem can be pruned.
+Only if the relaxed optimum turns out to be better than the best optimum so far it is necessary to look further into this sub problem and either test the left candidates or apply additional branching.
 

### 2014-05-04 paper.tex
restart strategies

@@
 A simple heuristic for branching would be to split the biggest domain into two equal sized intervals.
-	Bounding is then the step which transforms the Branch-and-reduce search into an optimization.
+
+	Bounding is then the part which transforms the Branch-and-reduce search into an optimization.
 To be able to do this an objective function is necessary.
 In the following I will assume that the objective function has to be minimized with a loss of generality.
+
 	Every Branch splits the total problem into two sub problems.
@@
 	\subsubsection{Restart Strategies}
+	Branching means always to apply certain heuristics and this heuristics may fail or produce bad results for the first execution.
+If it turns out that the search tree is build in a disadvantageous way a restart strategy may do the trick to improve the search.
+For such a restart the heuristic is improved, based on the things the search has learned to be a not so good solution during the first or first few attempts to build a search tree.
+	For practical application several variants of restart strategies were proposed \cite[p.~113]{handbookCP}
 

### 2014-05-04 paper.tex
conclusion

@@
 	\section{Conclusion}
-	\textit{conclusion}
+	Constraint programming is a very useful tool in practice, as it can solve abritrary combinatorial or optimization problems.
+Gecode, as representant of the various toolsets, implements a very hand interface to transform a mathematical description of a problem into code, which then can be used to solve the problem.
+In conclusion Gecode provides all the necessary tools, combines them with a nice interface, and is still very flexible concering optimzation for a certain area through the possibility to implement custom search strategies, based on some well studied and efficient algorithms.
 

### 2014-05-04 paper.tex
abstract

@@
 	\begin{abstract}	
-		\textit{last thing to write}	
+		Gecode is one of the various tools for constraint programming, a well-studied area which aims to solve arbitrary combinatorial and optimzation problems.
+The basement for Gecode are variables, their domains, constraints on these variables and strategies to search for a solution space or only single solutions.
 	\end{abstract}

### 2014-05-04 paper.tex
emphasized gecode

--- 2014-05-04 paper.tex
@@
 	\begin{abstract}	
-		\textit{last thing to write}	
+		\emph{Gecode} is one of the various tools for constraint programming, a well-studied area which aims to solve arbitrary combinatorial and optimzation problems.
+The basement for \emph{Gecode} are variables, their domains, constraints on these variables and strategies to search for a solution space or only single solutions.
 	\end{abstract}
@@
 
-	Now a solver for constraint programming like Gecode provides the necessary tools to define these constraints in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected search algorithm.
+	Now a solver for constraint programming like \emph{Gecode} provides the necessary tools to define these constraints in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected search algorithm.
 Actually this example contains only linear constraints and could therefore be solved through linear programming.
@@
 	\end{itemize}
-	In Gecode for example it is very easy to create for example a nonlinear constraint based on the integer variables a, b, c and d \cite[p.~120]{programmingGecode}:
+	In \emph{Gecode} for example it is very easy to create for example a nonlinear constraint based on the integer variables a, b, c and d \cite[p.~120]{programmingGecode}:
 	\begin{lstlisting}
@@
 	\end{equation}
-	and in Gecode by
+	and in \emph{Gecode} by
 	\begin{lstlisting}
@@
 	The actual implementation of constraint propagation differs again from one toolset to another, but the basic concepts are for sure the same.
-As the topic of this paper is the implementation in Gecode I will concentrate on this one.
+As the topic of this paper is the implementation in \emph{Gecode} I will concentrate on this one.
 
-	Constraints in Gecode are always evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the subtrees.
+	Constraints in \emph{Gecode} are always evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the subtrees.
 The problem here is to decide in which order the constraints should be considered, and it may even be necessary to evaluate some constraints more often.
@@
 
-	To implement this repeated evaluation of constraints in Gecode all constraints subscribe to the domains of the variables which are part of them.
+	To implement this repeated evaluation of constraints in \emph{Gecode} all constraints subscribe to the domains of the variables which are part of them.
 With this method the constraints get noticed whenever a variable domain is reduced and they can signal the system that they need to be evaluated once again.
@@
 In the example above the constraints were quite simple and therefore it won't make a big difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time.
-Unfortunately not all constraints are that easy to evaluate and for this case Gecode defines for every constraint a certain cost \cite[p.~275]{programmingGecode}.
+Unfortunately not all constraints are that easy to evaluate and for this case \emph{Gecode} defines for every constraint a certain cost \cite[p.~275]{programmingGecode}.
 Based on this cost the constraints are sorted into buckets which are then iterated from the constraints with lower costs to the ones with higher costs.
@@
 
-	In Gecode it is possible to define its own search strategies, but the user is supported at this point with some standard engines which can be used to build custom ones:
+	In \emph{Gecode} it is possible to define its own search strategies, but the user is supported at this point with some standard engines which can be used to build custom ones:
 	\begin{itemize}
@@
 	\section{Conclusion}
-	\textit{conclusion}
+	Constraint programming is a very useful tool in practice, as it can solve abritrary combinatorial or optimization problems.
+\emph{Gecode}, as representant of the various toolsets, implements a very hand interface to transform a mathematical description of a problem into code, which then can be used to solve the problem.
+In conclusion \emph{Gecode} provides all the necessary tools, combines them with a nice interface, and is still very flexible concering optimzation for a certain area through the possibility to implement custom search strategies, based on some well studied and efficient algorithms.
 

### 2014-05-04 paper.tex
corrections

@@
 	Constraint programming is a tool to solve mathematical problems through the formulation of constraints, which describe the desired solution.
-As typically a foraml and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint propagation, is not as helpful as an example, I would like to give you a short overview on the topic through the well known problem \emph{SEND MORE MONEY}.
+As usually a formal and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint programming, is not as helpful as an example, I would like to give you a short overview on the topic through the well known problem \emph{SEND MORE MONEY}.
 
@@
 As linear programming is only a subset of constraint programming the latter one is often the more useful one, because not every problem in the real world is linear or can be linearized in a useful way.
-Consequently constraint programming with its ability to consider constraints in there most general form can be used in various fields like scheduling of resources, computer-aided design and robotics \cite[p.~221]{trendsInCP}.
+Consequently constraint programming with its ability to consider constraints in their most general form can be used in various fields like scheduling of resources, computer-aided design and robotics \cite[p.~221]{trendsInCP}.
 
@@
 For integer and boolean variables the domain for each variable is reduced step by step during a search, either by constraints which rule out certain values or by trial and error.
-As floating variables do not have discrete values for their domains the procedere in this case is to reduce or split intervals based on constraints or to apply trial and error through splitting the intervals \cite[p.~571]{handbookCP}.
-I will come back to this topic later during the discussion of the branch-and-bound search.
+As floating variables do not have discrete values for their domains the procedere in this case is to reduce or split intervals based on constraints, or to apply trial and error through splitting the intervals \cite[p.~571]{handbookCP}.
+I will come back to this topic later during the discussion of branching strategies.
 
@@
 	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \refeq{eq:linearConstraint}.
-Which statements are possible to make varies than very much from the programming language in which the problem is implemented and the used toolset.
+Which statements are possible to make and how handy it is to do so varies very much from the programming language in which the problem is implemented and the used toolset.
 Typical constraints would be
@@
 	\end{itemize}
-	In \emph{Gecode} for example it is very easy to create for example a nonlinear constraint based on the integer variables a, b, c and d \cite[p.~120]{programmingGecode}:
+	In \emph{Gecode} it is very easy to create for instance a nonlinear constraint based on the integer variables $a$, $b$, $c$ and $d$ \cite[p.~120]{programmingGecode}:
 	\begin{lstlisting}
@@
 	\subsection{Search Algorithms}
-	In constraint programming search algorithm define the way the search tree is constructed, starting from the root node.
+	In constraint programming search algorithms define the way the search tree is constructed, starting from the root node.
 To improve the performance of the search usually at every newly created node a constraint propagation is executed to narrow the search space down.
@@
 
-	In \emph{Gecode} it is possible to define its own search strategies, but the user is supported at this point with some standard engines which can be used to build custom ones:
+	In \emph{Gecode} it is possible to define custom strategies, but the user is supported at this point with some standard engines which can be used to build custom ones:
 	\begin{itemize}
@@
 	For all branching strategies there is a big choice of freedom left, which can be used to implement some heuristics.
-This additional information is either gained during the search itself of pre defined by the user.
+This additional information is either gained during the search itself of predefined by the user.
 
@@
 The idea behind this search algorithm is that every possible combination of variables is examined.
-Obviously in its basic form it can not be applied to problems with inifinite domains; for problems with such variable domains it is necessary to use for example a Branch-and-Bound Search.
-Sometimes the Depth-First Search is also referred to as chronological backtrack search.
+Obviously in its basic form it can not be applied to problems with inifinite domains; for problems with such variable domains it is necessary to use for instance Branch-and-reduce.
+Sometimes the Depth-first search is also referred to as chronological backtrack search.
 
-	The starting point is the root node, from where one the algorithm is called recursively.
+	The starting point is the root node, from where on the algorithm is called recursively.
 At every node mainly two steps are executed: First a constraint propagation and then a branching, which divides the search space into two disjoint sub spaces, on which the same algorithm is applied again.
@@
 
-	By stopping the search after one feasible solution is found the algorithm the Depth-first search can often be accelerated.
+	By stopping the search after one feasible solution is found the algorithm can often be accelerated.
 Although this slightly modified version is not complete anymore, in practice it is in some areas sufficient to know just one valid solution.
@@
 	\subsubsection{Depth-First Search with Optimization}
-	If in the constraint programming is a certain objective is defined which should be either minimized or maximized a modified version of the Depth-first search can be applied.
+	If in the constraint programming a objective is defined which should be either minimized or maximized a modified version of the Depth-first search can be applied.
 The only difference is that it evalutes the objective function for every found solution concerning the constraints and stores during the search only the best one.
-To accelerate the search after one solution is already found it is possible to add an additional constraint which defines, that the new objective must be greater than the one related to the best solution till that point.
-Through this enhancement it is possible to prune even more sub trees and therefore to get the desired result faster.
+To accelerate the search it is possible to add an additional constraint which defines that the new objective must better than the best so far.
+Through this enhancement even more sub trees are pruned and the search is faster.
 
 	This algorithm is again a complete one, therefore it will always find the best solution if one exists.
-This algorithm does not inherit the bad memory characteristics from the Depth-First Search as only one solution is stored, but the search is still exhaustive and can be very slow, depending on the size of the variable domains.
+This algorithm does not inherit the bad memory characteristics from the Depth-first search as only one solution is stored, but the search is still exhaustive and can be very slow, depending on the size of the variable domains.
 
@@
 To be able to do this an objective function is necessary.
-In the following I will assume that the objective function has to be minimized with a loss of generality.
+In the following I will assume that the objective function has to be minimized without a loss of generality.
 
 	Every Branch splits the total problem into two sub problems.
-With some additional constraints, for example one variable must be an integer $x \in \mathbb{Z}$, it may happen that the real optimum of a sub problem is not the same one as if this additional constraint is considered.
+With some additional constraints, for example one variable must be an integer $x \in \mathbb{Z}$, it may happen that the optimum of a sub problem without this constraint is not the same one as if this additional constraint is considered.
 For finding the first optimum the trick is to branch in ways that a certain sub problem ends up with only a few possible candidates left, which then can be tested each.
@@
 	\subsubsection{Limited Discrepancy Search}
-	Limited Discrepancy Search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy} and it depends on some heurisitics for the tree, which are determined through branching strategies.
-The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive and second heurisitics work worse the farther away from the solution they are.
+	Limited Discrepancy Search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy}.
+It depends on some heurisitics for the tree, which are determined through branching strategies.
+The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive and second heurisitics work worse the farther away from the solution, therefore closer to the root, they are.
 Consequently, in a chronological backtrack search the worst decisions are the most expensive ones.
@@
 A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the deadend.
-How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is down wrong the backjumping can even be less efficient than chronological backtracking \cite[p.~100]{handbookCP}.
+How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is done wrong the backjumping can even be less efficient than chronological backtracking \cite[p.~100]{handbookCP}.
 
 	\subsubsection{Restart Strategies}
-	Branching means always to apply certain heuristics and this heuristics may fail or produce bad results for the first execution.
+	Branching means always to apply certain heuristics and these heuristics may fail or produce bad results for the first execution.
 If it turns out that the search tree is build in a disadvantageous way a restart strategy may do the trick to improve the search.
@@
 	Constraint programming is a very useful tool in practice, as it can solve abritrary combinatorial or optimization problems.
-\emph{Gecode}, as representant of the various toolsets, implements a very hand interface to transform a mathematical description of a problem into code, which then can be used to solve the problem.
-In conclusion \emph{Gecode} provides all the necessary tools, combines them with a nice interface, and is still very flexible concering optimzation for a certain area through the possibility to implement custom search strategies, based on some well studied and efficient algorithms.
+\emph{Gecode}, as representant of the various toolsets, implements a very handy interface to transform a mathematical description of a problem into code, which then can be used to solve the problem.
+In conclusion, \emph{Gecode} provides all the necessary tools, combines them with a nice interface, and is still very flexible concering optimzation for a certain area through the possibility to implement custom search strategies, based on some well studied and efficient algorithms.
 

### 2014-05-05 paper.tex
fixed references

@@
 
-\newcommand{\reffig}[1]{{[Fig.~\ref{#1}.]}}
-\newcommand{\refeq}[1]{{[Eq.~\ref{#1}.]}}
+\newcommand{\reffig}[1]{{figure \ref{#1}}}
+\newcommand{\refeq}[1]{{(\ref{#1})}}
 
@@
 	Hence there are boolean variables available it is also possible to use logical operators to combine other constraint types to logical expressions.
-To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap \reffig{fig:squares}.
+To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap.
+The related variables are defined in \reffig{fig:squares}.
 This problem can be expressed mathematically by \cite[p.
@@
 	\end{equation}
-	Without constraint propagation and branching first on $x$ this results in search tree \reffig{fig:bigTree}.
-Obviously the right parts of this tree can already be pruned already at the root by propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these subtrees \reffig{fig:smallTree}.
+	Without constraint propagation and branching first on $x$ the search trees turns out to be the one in \reffig{fig:bigTree}.
+Obviously the right parts of this tree can already be pruned already at the root by propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these subtrees like in \reffig{fig:smallTree}.
 

### 2014-05-05 paper.tex
removed C++-code and pseudo code

@@
 	\end{itemize}
-	In \emph{Gecode} it is very easy to create for instance a nonlinear constraint based on the integer variables $a$, $b$, $c$ and $d$ \cite[p.~120]{programmingGecode}:
-	\begin{lstlisting}
-rel(home, a+b*(c+d) == 0);
-	\end{lstlisting}
-	This single statement in C++ implements the constraint
-	\begin{equation}
-		a + b \cdot (c + d) == 0
-	\end{equation}
-	which is a quite handy way to do it.
 
@@
 The related variables are defined in \reffig{fig:squares}.
-This problem can be expressed mathematically by \cite[p.
+This problem can be expressed by \cite[p.
 101]{programmingGecode}
@@
 	\end{equation}
-	and in \emph{Gecode} by
-	\begin{lstlisting}
-rel(home,	(x1 + d1 <= x2) || 
-		(x2 + d2 <= x1) || 
-		(y1 + d1 <= y2) || 
-		(y2 + d2 <= y1));
-	\end{lstlisting}
-	which is again quite obvious to a reader with some experience in a C-like programming language.
 
@@
 
-	\subsubsection{Depth-First Search}
-	\begin{algorithm}
-		\begin{algorithmic}
-			\Function{DFS}{node}
-				\If{$Goal(node)$}
-					\State \Return $node$
-				\EndIf
-				\State $s \gets Successors(node)$
-				\If{$Null(s)$}
-					\State \Return $nil$
-				\EndIf
-				\State $result \gets DFS(Left(s))$
-				\If{$result \ne nil$}
-					\State \Return $result$
-				\Else
-					\State \Return $DFS(Right(s))$
-				\EndIf
-			\EndFunction
-		\end{algorithmic}
-		\caption{Depth-first search in a binary search tree for one feasible solution, without constraint propagation}
-	\end{algorithm}
-
 	The Depth-first search is a good point to start, as it was also historical the first developed one.
@@
 
-	\begin{algorithm}
-		\begin{algorithmic}
-			\Function{$LDSProbe$}{$node$, $k$}
-				\If{$Goal(node)$}
-				    \State \Return $node$
-				\EndIf
-				\State $s \gets Successors(node)$
-				\If{$Null(s)$}
-					\State \Return $nil$
-				\EndIf
-				\If{$k = 0$}
-					\State \Return $LDSProbe(Left(s), 0)$
-				\Else
-					\State $result \gets LDSProbe(Right(s), k - 1)$
-					\If{$result \ne nil$}
-						\State \Return $result$
-					\EndIf
-					\State \Return $LDSProbe(Left(s), k)$
-				\EndIf
-			\EndFunction
-		\end{algorithmic}
-		\caption{Probe of Limited Discrepancy Search for a binary search tree from \cite{limitedDiscrepancy}, without constraint propagation}
-	\end{algorithm}
-
-	\begin{algorithm}
-		\begin{algorithmic}			
-			\Function{$LDS$}{$node$}
-				\For{$x \gets 0\ \text{to maximum depth}$}
-					\State $result \gets LDSProbe(node, x)$
-					\If{$result \ne nil$}
-						\State \Return $result$
-					\EndIf
-				\EndFor
-			\EndFunction
-		\end{algorithmic}
-		\caption{Limited Discrepancy Search for a binary search tree from \cite{limitedDiscrepancy}, without constraint propagation}
-	\end{algorithm}
-
 	\subsubsection{Backjumping}

### 2014-05-05 paper.tex
removed C++-code and pseudo code

@@
 	\section{Introduction}
-	Constraint programming is a tool to solve mathematical problems through the formulation of constraints, which describe the desired solution.
-As usually a formal and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint programming, is not as helpful as an example, I would like to give you a short overview on the topic through the well known problem \emph{SEND MORE MONEY}.
+	Constraint programming is a method to solve mathematical problems through the formulation of constraints, which describe the desired solution.
+As a formal and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint programming, is not as helpful as an example, I would like to give you a short overview on the topic through the well known problem \emph{SEND MORE MONEY}.
 

### 2014-05-05 paper.tex
corrections Michael

@@
 	\begin{abstract}	
-		\emph{Gecode} is one of the various tools for constraint programming, a well-studied area which aims to solve arbitrary combinatorial and optimzation problems.
-The basement for \emph{Gecode} are variables, their domains, constraints on these variables and strategies to search for a solution space or only single solutions.
+		NEXT TRY
 	\end{abstract}
@@
 	Constraint programming is a method to solve mathematical problems through the formulation of constraints, which describe the desired solution.
-As a formal and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint programming, is not as helpful as an example, I would like to give you a short overview on the topic through the well known problem \emph{SEND MORE MONEY}.
+As a formal and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint programming, is not as helpful as an example, I would like to give you a short overview on the topic through the problem called \emph{SEND MORE MONEY}.
 
@@
 	\[SEND + MORE = MONEY\]
-	where every character is a variable and the position in the word describes the weight.
+	where every character is a variable and the position in the word describes the weight, like it can be seen in \refeq{eq:linearConstraint}.
 	\begin{equation}
@@
 
-	To narrow the solution space a little bit down I state additional constraints:
+	To narrow the solution space down I state additional constraints:
 	\begin{enumerate}
@@
 		\begin{equation}
+		\label{eq:allDifferent}
 		\begin{split}
@@
 		\begin{equation}
+			\label{eq:trailingZero}
 			S \ne 0 \land M \ne 0
@@
 
-	Now a solver for constraint programming like \emph{Gecode} provides the necessary tools to define these constraints in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected search algorithm.
-Actually this example contains only linear constraints and could therefore be solved through linear programming.
-As linear programming is only a subset of constraint programming the latter one is often the more useful one, because not every problem in the real world is linear or can be linearized in a useful way.
-Consequently constraint programming with its ability to consider constraints in their most general form can be used in various fields like scheduling of resources, computer-aided design and robotics \cite[p.~221]{trendsInCP}.
+	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:linearConstraint}, \refeq{eq:allDifferent} and \refeq{eq:trailingZero} in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected algorithm.
+This example contains only linear constraints and could therefore be solved through linear programming REFERENCE MISSING.
+As linear programming is only a subset of constraint programming the latter one is the more general one with more practical applications; not every problem in the real world is linear or can be linearized in a useful way.
+Consequently, constraint programming with its ability to consider constraints in their most general form can be used in various fields like scheduling of resources, computer-aided design and robotics \cite[p.~221]{trendsInCP}.
 
@@
 	\subsection{Variable Domains}
-	In constraint programming to a variable a set of possible values is connected, the so called domain.
-This domain can be a finite or inifite set of integers, true or false and intervals.
+	In constraint programming a variable is connected to a set of possible values, the so called domain.
 Depending on the domain a variable is called an integer, boolean or floating variable.
@@
 	\begin{itemize}
-		\item Integer: $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$, \dots
+		\item Integer: finite or inifnite sets like $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$, \dots
 		\item Bool: $x \in \{\text{true}, \text{false}\}$
-		\item Float: $x \in [-4, 3]$, $x \in [-10, 8] \cup [10, 12]$, \dots
+		\item Float: intervals or unions of intervals like $x \in [-4, 3]$, $x \in [-10, 8] \cup [10, 12]$, \dots
 	\end{itemize}
-	This distinction is important as especially for floating variables the classical approach to constraint programming, a backtrack search, can not be applied directly.
-For integer and boolean variables the domain for each variable is reduced step by step during a search, either by constraints which rule out certain values or by trial and error.
-As floating variables do not have discrete values for their domains the procedere in this case is to reduce or split intervals based on constraints, or to apply trial and error through splitting the intervals \cite[p.~571]{handbookCP}.
-I will come back to this topic later during the discussion of branching strategies.
+	This distinction is fundamental as especially for floating variables the classical combinatorial approach to constraint programming is impossible.
 
@@
 
-	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \refeq{eq:linearConstraint}.
+	As constraint programming is the generalization of linear programming it is at least possible do define linear constraints like \refeq{eq:linearConstraint}.
 Which statements are possible to make and how handy it is to do so varies very much from the programming language in which the problem is implemented and the used toolset.
-Typical constraints would be
+Typical constraints are the following
 	\begin{itemize}
@@
 
-	Hence there are boolean variables available it is also possible to use logical operators to combine other constraint types to logical expressions.
+	Different constraint types like equalities and inequalities can be combined through logical operators to boolean expressions.
 To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap.
@@
 
+	\subsection{Search Tree}
+	MISSING
+
 	\subsection{Constraint Propagation}
-	Constraint propagation is a key concept in constraint programming as it allows to speed-up the search significantly.
+	Constraint propagation is a key concept in constraint programming as it reduces the size of the search space significantly.
 The term propagation refers in this area to the reduction of domains of variables which are infeasible, considering other constraints and affected variable domains.
@@
 
-	The actual implementation of constraint propagation differs again from one toolset to another, but the basic concepts are for sure the same.
+	The actual implementation of constraint propagation differs again from one toolset to another, but the basic concepts are for sure the same REFERENZ FEHLT.
 As the topic of this paper is the implementation in \emph{Gecode} I will concentrate on this one.
@@
 	\end{equation}
-	Actually this behaviour is very nice to have as it narrows efficiently the search space down.
 
@@
 
-	The next difficult decision is which constraint should be evaluated first.
+	The next decision to be made is which constraint should be evaluated first.
 In the example above the constraints were quite simple and therefore it won't make a big difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time.
@@
 Based on these two steps then either one, the best or no solution is found.
-What the actual result in the end is depends obviously on the search algorithm.
+What the actual result in the end is depends on the type of search algorithm.
 
@@
 
-	In the following part I will discuss these searches and some additional, well known techniques, which can be applied during searches in constraint programming.
+	In the following part I will discuss these searches and additional techniques, which can be applied during searches in constraint programming.
 
 	\subsubsection{Branching Strategies}
-	If at one node it is not possible to empty one variable domain or to reduce all variable domains to the size one it is necessary to make a decision: How should be the next subtree be build?
-The question is answered by the selected branching strategy, which depends on the variable domain.
-For Integer (and Bool) Variables there are three possible strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the subtrees.
-All the strategies have in common that they create additional temporary constraints, which are only valid for the specific subtree.
-The temporay addition of constraints is called \emph{posting}.
+	If at one node it is not possible to empty one variable domain or to reduce all variable domains to the size one it is necessary to make a decision: How should the next subtree be constructed?
+The question is answered by the selected branching strategy.
+No matter which one is selected, they all have in common that they create additional, temporary constraints, which are only applied to the specific subtrees.
+This step is called \emph{posting} a constraint on a branch.
+
+	The branching strategy depends on the variable domain: For Integer (and Boolean) variables there are three common strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the subtrees.
 	\begin{itemize}
@@
 
-	Obviously for variables with a Float domain the Enumeration and Binary choice points are not feasible, but the last one, domain splitting, is typically used.
-As termination criteria for a Float it is often necessary to define a epsilon which describes the interval length, at which the solver states to have found a solution.
+	As for variables with a Float domain the Enumeration and Binary choice points are not feasible, the last one, domain splitting, is typically used.
+As termination criteria for a Float it is often necessary to define a certain epsilon $\epsilon \in \mathbb{R}_{>0}$ which describes the interval length, at which the solver states to have found a solution.
 
-	For all branching strategies there is a big choice of freedom left, which can be used to implement some heuristics.
+	For all branching strategies there is a choice of freedom left, which can be used to implement some heuristics.
 This additional information is either gained during the search itself of predefined by the user.

### 2014-05-06 paper.tex
addd bibitem
added reference

@@
 	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:linearConstraint}, \refeq{eq:allDifferent} and \refeq{eq:trailingZero} in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected algorithm.
-This example contains only linear constraints and could therefore be solved through linear programming REFERENCE MISSING.
+This example contains only linear constraints and could therefore be solved through linear programming \cite[1]{linearProgramming}.
 As linear programming is only a subset of constraint programming the latter one is the more general one with more practical applications; not every problem in the real world is linear or can be linearized in a useful way.
@@
 		W.~D.~Harvey, M.~L.~Ginsberg, \emph{Limited Discrepancy Search}, Morgan Kaufmann, 1995
+		\bibitem{linearProgramming}
+		L.~Cooper and D.~Steinberg, \emph{Methods and Applications of Linear Programming}, W.~B.~Saunder~Company, 1974, ISBN: 0-7216-2694-7
 	\end{thebibliography}

### 2014-05-06 paper.tex
corrections Michael

@@
 	\subsubsection{Depth-First Search with Optimization}
-	If in the constraint programming a objective is defined which should be either minimized or maximized a modified version of the Depth-first search can be applied.
+	If in the constraint programming an objective is defined which should be either minimized or maximized a modified version of the Depth-first search can be applied.
 The only difference is that it evalutes the objective function for every found solution concerning the constraints and stores during the search only the best one.
@@
 	Branch-and-bound refers actually to two parts, a Branch-and-reduce search and bounding.
-Branch-and-reduce itself is the continous constraint propagation to reduce the variable domains combined with branching when there is no further reduction by constraint propagation possible.
-A simple heuristic for branching would be to split the biggest domain into two equal sized intervals.
+If there is no further reduction of variable domains possible the step of Branch-and-reduce can be applied, which means applying a branching strategy to split the search space into two disjoints sets.
+One example for a heuristic for branching would be to split the biggest domain into two equal sized intervals.
 
@@
 To be able to do this an objective function is necessary.
-In the following I will assume that the objective function has to be minimized without a loss of generality.
+In the following discussion on bounding I will assume that the objective function has to be minimized without a loss of generality.
 
@@
 It depends on some heurisitics for the tree, which are determined through branching strategies.
-The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive and second heurisitics work worse the farther away from the solution, therefore closer to the root, they are.
+The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive.
+Second, heuristics work worse the farther away from the solution they are applied, therefore decisions close to the root node are possible to fail.
 Consequently, in a chronological backtrack search the worst decisions are the most expensive ones.
@@
 	Constraint programming is a very useful tool in practice, as it can solve abritrary combinatorial or optimization problems.
-\emph{Gecode}, as representant of the various toolsets, implements a very handy interface to transform a mathematical description of a problem into code, which then can be used to solve the problem.
+\emph{Gecode}, as representant of the various toolsets, implements a handy interface to transform a mathematical description of a problem into code, which then can be used to solve the problem.
 In conclusion, \emph{Gecode} provides all the necessary tools, combines them with a nice interface, and is still very flexible concering optimzation for a certain area through the possibility to implement custom search strategies, based on some well studied and efficient algorithms.

### 2014-05-06 paper.tex
added introductionary part to the section on search algorithms

@@
 
-	In the following part I will discuss these searches and additional techniques, which can be applied during searches in constraint programming.
+	In the following part I will discuss branching strategies, the Depth-first search and one of its variants, the Branch-and-bound search, the Limited-Discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
 
@@
 
+	\subsubsection{Depth-First Search}
 	The Depth-first search is a good point to start, as it was also historical the first developed one.

### 2014-05-06 paper.tex
minor improvements

@@
 
-	\subsection{Search Tree}
+	\subsection{Search Trees}
 	MISSING
@@
 The term propagation refers in this area to the reduction of domains of variables which are infeasible, considering other constraints and affected variable domains.
-As the size of the variable domain is related to possibilities which have to be checked by a search, the constraint propagation is able to prune infeasible subtrees.
-Therefore the whole search process is accelerated.
+As the size of the variable domain is related to possibilities which have to be checked by a search, the constraint propagation is able to prune complete infeasible subtrees.
+Therefore the whole search process can be accelerated through constraint propagation.
 

### 2014-05-06 paper.tex
replaced example

@@
 
-	Consider the problem
-	\[SEND + MORE = MONEY\]
-	where every character is a variable and the position in the word describes the weight, like it can be seen in \refeq{eq:linearConstraint}.
+	Consider the problem of choosing how many components of three different types should be placed on a chip where variables $x$, $y$ and $z$ represent the number of components for each type.
+As the size of the chip is limited the total area is constrainted by
 	\begin{equation}
-	\label{eq:linearConstraint}
-	\begin{split}
-		& S \cdot 10^3 + E \cdot 10^2 + N \cdot 10^1 + D \cdot 10^0 + \\ 
-		& M \cdot 10^3 + O \cdot 10^2 + R \cdot 10^1 + E \cdot 10^0 = \\ 
-		& M \cdot 10^4 + O \cdot 10^3 + N \cdot 10^2 + E \cdot 10^1 + Y \cdot 10^0
-	\end{split}
+		10x + 15y + 5z \le 1000
+		\label{eq:areaConstraint}
+	\end{equation}
+	with different weights for each type. Additionally, the types have different costs and the total cost for a chip is limited:
+	\begin{equation}
+		3x + y^2 + 2z \le 500
+		\label{eq:costConstraint}
+	\end{equation}
+	The target for this problem could be maximization of the price at which the chip can be sold:
+	\begin{equation}
+		\max_{x, y, z} 5x + 6\ln(y) + z
+		\label{eq:objective}
 	\end{equation}
 
-	To narrow the solution space down I state additional constraints:
-	\begin{enumerate}
-	\item All variables must have different values.
-		\begin{equation}
-		\label{eq:allDifferent}
-		\begin{split}
-			S \ne E \land S \ne N \land & S \ne D \land \dots \land \\
-			E \ne N \land & E \ne D \land \dots \land \\
-			& N \ne D \land \dots
-		\end{split}
-		\end{equation}	
-	\item There must not be a trailing zero. Combined with the so-called all-different constraint \cite{allDifferent} from above this results in
-		\begin{equation}
-			\label{eq:trailingZero}
-			S \ne 0 \land M \ne 0
-		\end{equation}
-	\end{enumerate}
-
-	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:linearConstraint}, \refeq{eq:allDifferent} and \refeq{eq:trailingZero} in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected algorithm.
-This example contains only linear constraints and could therefore be solved through linear programming \cite[1]{linearProgramming}.
+	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:areaConstraint}, \refeq{eq:costConstraint} and \refeq{eq:objective} in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected algorithm.
+If there would be only linear constraints like \refeq{eq:areaConstraint} it would be possible to solve this problem through linear programming \cite[1]{linearProgramming}.
 As linear programming is only a subset of constraint programming the latter one is the more general one with more practical applications; not every problem in the real world is linear or can be linearized in a useful way.

### 2014-05-06 paper.tex
improved example of two squares

@@
 
-	As constraint programming is the generalization of linear programming it is at least possible do define linear constraints like \refeq{eq:linearConstraint}.
+	As constraint programming is the generalization of linear programming it is at least possible do define linear constraints like \refeq{eq:areaConstraint}.
 Which statements are possible to make and how handy it is to do so varies very much from the programming language in which the problem is implemented and the used toolset.
@@
 	Different constraint types like equalities and inequalities can be combined through logical operators to boolean expressions.
-To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap.
-The related variables are defined in \reffig{fig:squares}.
-This problem can be expressed by \cite[p.
-101]{programmingGecode}
+To show the usage of this  I want to come back to the example of the introduction, but this time with a different focus.
+Consider the number of components are already chosen and this time the task is to place the components on the chip.
+To simplify things I will only take two squares, which then should be placed such that they do not overlap.
+Of course, in a real world example first of all not only squares, but a lot more objects would have to be considered.
+Second, the objects will not be only squares and finally there would have to be a lot more constraints to be considered like minimum distances, a minimzation of the area usage, routing and so forth.
+
+	The related variables for this example are defined in \reffig{fig:squares}. The problem can be expressed by \cite[p. 101]{programmingGecode}
 	\begin{equation}
@@
 	\end{equation}
+	The task for the constraint programming tool at this point is then to find feasible solutions which fulfill the constraints.
 

### 2014-05-06 paper.tex
made squares to the introduction example

--- 2014-05-06 paper.tex
@@
 	Constraint programming is a method to solve mathematical problems through the formulation of constraints, which describe the desired solution.
-As a formal and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint programming, is not as helpful as an example, I would like to give you a short overview on the topic through the problem called \emph{SEND MORE MONEY}.
+As a formal and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint programming, is not as helpful as an example, I would like to give you a short overview on the topic through the problem of the placement of two squares such that they do not overlap.
 
-	Consider the problem of choosing how many components of three different types should be placed on a chip where variables $x$, $y$ and $z$ represent the number of components for each type.
-As the size of the chip is limited the total area is constrainted by
+	\begin{figure}
+		\center
+		\begin{tikzpicture}
+			\draw (0, 0) rectangle (1, 1);
+			\node at (0, -0.3) {($x_1$, $y_1$)};
+			\draw[arrows=<->] (1.2, 0) -- (1.2, 1);
+			\node at (1.5, 0.5) {$d_1$};
+			\draw (1.5, 1.5) rectangle(2, 2);
+			\node at (1.5, 1.2) {($x_2$, $y_2$)};
+			\draw[arrows=<->] (2.2, 1.5) -- (2.2, 2);
+			\node at (2.5, 1.75) {$d_2$};
+		\end{tikzpicture}
+		\caption{two squares which should not overlap}
+		\label{fig:squares}
+	\end{figure}
+
+	The related variables for this example are defined in \reffig{fig:squares}. The problem can be expressed by \cite[p. 101]{programmingGecode}
 	\begin{equation}
-		10x + 15y + 5z \le 1000
-		\label{eq:areaConstraint}
+	\begin{split}
+		x_1 + d_1 \le x_2 & \lor x_2 + d_2 \le x_1 \lor \\
+		y_1 + d_1 \le y_2 & \lor y_2 + d_2 \le y_1
+	\end{split}
+	\label{eq:squares}
 	\end{equation}
-	with different weights for each type. Additionally, the types have different costs and the total cost for a chip is limited:
-	\begin{equation}
-		3x + y^2 + 2z \le 500
-		\label{eq:costConstraint}
-	\end{equation}
-	The target for this problem could be maximization of the price at which the chip can be sold:
-	\begin{equation}
-		\max_{x, y, z} 5x + 6\ln(y) + z
-		\label{eq:objective}
-	\end{equation}
+	The task for the constraint programming tool at this point is then to find feasible solutions which fulfill the constraints.
 
-	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:areaConstraint}, \refeq{eq:costConstraint} and \refeq{eq:objective} in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected algorithm.
-If there would be only linear constraints like \refeq{eq:areaConstraint} it would be possible to solve this problem through linear programming \cite[1]{linearProgramming}.
+	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:squares} in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected algorithm.
+If there would be only linear constraints like \refeq{eq:squares} it would be possible to solve this problem through linear programming \cite[1]{linearProgramming}.
 As linear programming is only a subset of constraint programming the latter one is the more general one with more practical applications; not every problem in the real world is linear or can be linearized in a useful way.
@@
 
-	\begin{figure}
-	\center
-	\begin{tikzpicture}
-		\draw (0, 0) rectangle (1, 1);
-		\node at (0, -0.3) {($x_1$, $y_1$)};
-		\draw[arrows=<->] (1.2, 0) -- (1.2, 1);
-		\node at (1.5, 0.5) {$d_1$};
-		\draw (1.5, 1.5) rectangle(2, 2);
-		\node at (1.5, 1.2) {($x_2$, $y_2$)};
-		\draw[arrows=<->] (2.2, 1.5) -- (2.2, 2);
-		\node at (2.5, 1.75) {$d_2$};
-	\end{tikzpicture}
-	\caption{two squares which should not overlap}
-	\label{fig:squares}
-	\end{figure}
-
-	As constraint programming is the generalization of linear programming it is at least possible do define linear constraints like \refeq{eq:linearConstraint}.
-Which statements are possible to make and how handy it is to do so varies very much from the programming language in which the problem is implemented and the used toolset.
-Typical constraints are the following
+	As constraint programming is the generalization of linear programming it is at least possible do define linear constraints like \refeq{eq:squares}.
+Which statements are possible to make and how handy it is to do so varies from the programming language in which the problem is implemented and the used toolset.
+Typical constraints are
 	\begin{itemize}
@@
 	\end{itemize}
-
-	Different constraint types like equalities and inequalities can be combined through logical operators to boolean expressions.
-To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap.
-The related variables are defined in \reffig{fig:squares}.
-This problem can be expressed by \cite[p.
-101]{programmingGecode}
-	\begin{equation}
-	\begin{split}
-		x_1 + d_1 \le x_2 & \lor x_2 + d_2 \le x_1 \lor \\
-		y_1 + d_1 \le y_2 & \lor y_2 + d_2 \le y_1
-	\end{split}
-	\end{equation}
 

### 2014-05-06 paper.tex
avoided statement without reference

@@
 	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:squares} in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected algorithm.
-If there would be only linear constraints like \refeq{eq:squares} it would be possible to solve this problem through linear programming \cite[1]{linearProgramming}.
+If there would be only linear constraints like it would be possible to solve this problem through linear programming \cite[1]{linearProgramming}.
 As linear programming is only a subset of constraint programming the latter one is the more general one with more practical applications; not every problem in the real world is linear or can be linearized in a useful way.
@@
 
-	As constraint programming is the generalization of linear programming it is at least possible do define linear constraints like \refeq{eq:squares}.
+	As constraint programming is the generalization of linear programming it is at least possible to define linear constraints.
 Which statements are possible to make and how handy it is to do so varies from the programming language in which the problem is implemented and the used toolset.
@@
 
-	The actual implementation of constraint propagation differs again from one toolset to another, but the basic concepts are for sure the same REFERENZ FEHLT.
+	The actual implementation of constraint propagation differs again from one toolset to another, but theory underneath is the same.
 As the topic of this paper is the implementation in \emph{Gecode} I will concentrate on this one.

### 2014-05-06 paper.tex
added search tree

@@
 	\subsection{Search Trees}
-	MISSING
+
+	\begin{figure}
+		\center
+		\begin{tikzpicture}
+			\tikzstyle{arrow}=[draw, -latex]
+			\node[rectangle,draw] at (0, 2) (levelOne) {$x \in \mathbb{R}$};
+
+			\node[rectangle,draw] at (-3, 1) (levelTwoElementOne) {$x \le 0$};
+			\coordinate[left=0.2cm of levelOne.south] (levelOnePointOne);
+	  		\path[arrow] (levelTwoElementOne) -- (levelOnePointOne);
+
+			\node[rectangle,draw] at (0.5, 0.7) (levelTwoElementTwo) {$x > 0$};
+			\coordinate[right=0.2cm of levelOne.south] (levelOnePointTwo);
+	  		\path[arrow] (levelTwoElementTwo) -- (levelOnePointTwo);
+
+			\node[rectangle,draw] at (-4, 0.2) (levelThreeElementOne) {$x < -2$};
+			\coordinate[left=0.2cm of levelTwoElementOne.south] (levelTwoElementOnePointOne);
+	  		\path[arrow] (levelThreeElementOne) -- (levelTwoElementOnePointOne);
+
+			\node[rectangle,draw] at (-2, 0.2) (levelThreeElementTwo) {$x \ge -2$};
+			\coordinate[right=0.2cm of levelTwoElementOne.south] (levelTwoElementOnePointTwo);
+	  		\path[arrow] (levelThreeElementTwo) -- (levelTwoElementOnePointTwo);
+
+			\node[rectangle,draw] at (-1.5, -0.5) (levelThreeElementThree) {$x = 1$};
+			\coordinate[left=0.2cm of levelTwoElementTwo.south] (levelTwoElementTwoPointOne);
+	  		\path[arrow] (levelThreeElementThree) -- (levelTwoElementTwoPointOne);
+
+			\node[rectangle,draw] at (0, -0.5) (levelThreeElementFour) {$x = 2$};
+			\coordinate[right=0cm of levelTwoElementTwo.south] (levelTwoElementTwoPointTwo);
+	  		\path[arrow] (levelThreeElementFour) -- (levelTwoElementTwoPointTwo);
+
+			\node[rectangle,draw] at (2, -0.5) (levelThreeElementFive) {$x \ne 2 \land x \ne 1$};
+			\coordinate[right=0.2cm of levelTwoElementTwo.south] (levelTwoElementTwoPointThree);
+	  		\path[arrow] (levelThreeElementFive) -- (levelTwoElementTwoPointThree);
+		\end{tikzpicture}
+		\caption{search tree on one variable $x$}
+		\label{fig:searchTree}
+	\end{figure}
+
+	In constraint programming search trees, like the one in \reffig{fig:searchTree}, describe the partition of the search space and their nodes have names, depending on position in the tree.
+The root node is the one from where the search is started and all variable domains are still complete.
+This one is also the first one with a branch, where for the first time the search space is split up into disjoint sets.
+This approach is similiar to the divide-and-conquer technique as it reduces a complex problem into two or more parts with less complexity.
+The splitting, which adds additional constraints on the variables, is then recursively done with different heuristics to partition the search space into smaller parts.
+If one of the parts is small enough to decide if it is feasible the node can be called a either a deadend or a solution.
+To end up as the dead end there must be no feasible solutions left in the search space at this node.
+To turn the node into a solution it is necessary that all members of the reduced search space are valid considering the constraints.
 

### 2014-05-06 paper.tex
added alldifferent constraint

@@
 	\end{itemize}
+
+	Different constraint types can also be combined, for example inequalities or equalities through logic operators to boolean expressions.
+Some of these constraints may acquire special attention during the search for solutions to improve the execution reason.
+For this reason a lot of work has been done for example on the alldifferent constraint \cite{allDifferent}, which would be such a combined constraint.
+	\begin{equation}
+		\forall x_i, x_j \in D, i \ne j: x_i \ne x_j, 
+		\label{eq:alldifferent}
+	\end{equation}
 

### 2014-05-08 paper.tex
improved part on branch-and-bound

@@
 \usetikzlibrary{positioning}
+\usetikzlibrary{patterns}
 \usepackage{listings}
@@
 	\subsubsection{Branch-and-Bound Search}
-	Branch-and-bound refers actually to two parts, a Branch-and-reduce search and bounding.
-If there is no further reduction of variable domains possible the step of Branch-and-reduce can be applied, which means applying a branching strategy to split the search space into two disjoints sets.
-One example for a heuristic for branching would be to split the biggest domain into two equal sized intervals.
+	A branch-and-bound search is an algorithm for optimization of an objective function under certain constraints and only discrete possible candidates for a solution.
+To fulfill the constraints the first step, branch-and-reduce is applied, and the second one, bounding, is used to prune whole areas with non-optimal solutions.
 
-	Bounding is then the part which transforms the Branch-and-reduce search into an optimization.
-To be able to do this an objective function is necessary.
-In the following discussion on bounding I will assume that the objective function has to be minimized without a loss of generality.
+	\begin{figure}
+		\begin{tikzpicture}
+			\coordinate (a) at (0, 0);
+			\coordinate (b) at (4, 0);
+			\coordinate (c) at (5, 3);
+			\coordinate (d) at (3, 6);
+			\coordinate (e) at (-1, 2);
+			\coordinate (f) at (2, 0);
+			\coordinate (g) at (intersection of e--c and f--d);
 
-	Every Branch splits the total problem into two sub problems.
-With some additional constraints, for example one variable must be an integer $x \in \mathbb{Z}$, it may happen that the optimum of a sub problem without this constraint is not the same one as if this additional constraint is considered.
-For finding the first optimum the trick is to branch in ways that a certain sub problem ends up with only a few possible candidates left, which then can be tested each.
-With this first optimum the search algorithm has then a so far best optimum which can be used as upper bound for the optimum.
-If for one sub problem the relaxed optimum, the optimum without for example the constraint of integrity, is worse than this upper bound the sub problem can be pruned.
-Only if the relaxed optimum turns out to be better than the best optimum so far it is necessary to look further into this sub problem and either test the left candidates or apply additional branching.
+			\draw[very thick, fill=blue, fill opacity=0.1] (a) -- (b) -- (c) -- (d) -- (e) -- (a);
+			\draw[pattern=north west lines, pattern color=gray] (a) -- (b) -- (c) -- (e) -- (a);
+			\draw[pattern=north east lines, pattern color=gray] (f) -- (g) -- (c) -- (b) -- (f);
+		\end{tikzpicture}	
+		\center
+		\caption{Branch-and-reduce applied to a search space}
+		\label{fig:branchAndReduce}
+	\end{figure}
+
+	As the basic concept of the finding solutions is the enumeration of possible candidates it is necessary to seperate the search space in a useful way, which is the task for the branch-and-reduce step.
+Actually this step is very similar to the concept of branching, it also depends on a certain heuristic to divide the search space into two or even more disjoint parts.
+This heuristic can either be selected by the user or be an general one, like splitting the biggest variable domain into two equal sized parts.
+If the basic search space contains a big amount of possible candidate solution one reduction may not be enough and therefore this step is applied recursively to the newly create subproblems.
+This process can be stopped if the number of solutions is reduced to an amount, where it is possible to enumerate all of them and check if they are optimal.
+A possible division into subproblems can be seen in \reffig{fig:branchAndReduce}, where for legibility always only one of the subproblems is divided further.
+
+	The next step after branch-and-reduce is bounding, which is able to prune whole sub problems.
+In the following part I will consider the minimization of an objective function without a loss of generality.
+For the bounding it is first necessary to initialize a upper bound with a value, for example with $\infty$.
+This upper bound is updated during the process every time a new optimal solution is found to this new optimum.
+If a new possible solution is found which is above this upper bound the algorithm can prune it, as only the minimum solution is of interest.
+The actually pruning of subproblems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
+If this relaxed optimum for a certain subproblem is greater than the current upper bound it is possible to prune this sub problem, as it can not contain a better solution.
+
+	\begin{figure}
+		\begin{tikzpicture}
+			\coordinate (a) at (0, 0);
+			\coordinate (b) at (4, 0);
+			\coordinate (c) at (5, 3);
+			\coordinate (d) at (3, 6);
+			\coordinate (e) at (-1, 2);
+			\coordinate (f) at (2, 0);
+			\coordinate (g) at (intersection of e--c and f--d);
+
+			\begin{scope}
+				\clip (a) -- (b) -- (c) -- (d) -- (e) -- (a);
+				\foreach \x in {-2,-1.5,...,5}
+				{
+	      			\foreach \y in {0,0.5,...,6}
+	      			{
+	        			\node[draw,circle,inner sep=0.2pt,fill,gray] at (\x,\y) {};
+	      			}
+	    		}
+    		\end{scope}
+
+			\draw[very thick] (a) -- (b) -- (c) -- (d) -- (e) -- (a);
+			\draw (a) -- (b) -- (c) -- (e) -- (a);
+			\draw (f) -- (g) -- (c) -- (b) -- (f);
+			\node at (2.5, 4) {$P_0$};
+			\node at (1, 1) {$P_1$};
+			\node at (3.5, 1.5) {$P_2$};
+		\end{tikzpicture}	
+		\center
+		\caption{Possible candidates for a solution and the related search space}
+		\label{fig:bounding}
+	\end{figure}
+
+	Consider for this \reffig{fig:bounding}, where already a certain optimum $f_1$ was found in subproblem $P_1$. If now the relaxed optimum $f_{P0}$ is worse than this optimum
+	\begin{equation}
+		f_{P0} > f_1
+	\end{equation}
+	the whole subproblem $P_{0}$ can be pruned and it is not necessary to enumerate the solution candidates in it (or to apply split it further up).
+
+	Combined with appropriate branching the bounding can prune several candiates at once, without the necessity to enumerate them. Therefore bounding is able to improve the performance of the search.
 

### 2014-05-08 paper.tex
improved part on branch-and-bound
centered figures
reduced two figures to only one

@@
 	\begin{figure}
-		\begin{tikzpicture}
-			\coordinate (a) at (0, 0);
-			\coordinate (b) at (4, 0);
-			\coordinate (c) at (5, 3);
-			\coordinate (d) at (3, 6);
-			\coordinate (e) at (-1, 2);
-			\coordinate (f) at (2, 0);
-			\coordinate (g) at (intersection of e--c and f--d);
-
-			\draw[very thick, fill=blue, fill opacity=0.1] (a) -- (b) -- (c) -- (d) -- (e) -- (a);
-			\draw[pattern=north west lines, pattern color=gray] (a) -- (b) -- (c) -- (e) -- (a);
-			\draw[pattern=north east lines, pattern color=gray] (f) -- (g) -- (c) -- (b) -- (f);
-		\end{tikzpicture}	
 		\center
-		\caption{Branch-and-reduce applied to a search space}
-		\label{fig:branchAndReduce}
-	\end{figure}
-
-	As the basic concept of the finding solutions is the enumeration of possible candidates it is necessary to seperate the search space in a useful way, which is the task for the branch-and-reduce step.
-Actually this step is very similar to the concept of branching, it also depends on a certain heuristic to divide the search space into two or even more disjoint parts.
-This heuristic can either be selected by the user or be an general one, like splitting the biggest variable domain into two equal sized parts.
-If the basic search space contains a big amount of possible candidate solution one reduction may not be enough and therefore this step is applied recursively to the newly create subproblems.
-This process can be stopped if the number of solutions is reduced to an amount, where it is possible to enumerate all of them and check if they are optimal.
-A possible division into subproblems can be seen in \reffig{fig:branchAndReduce}, where for legibility always only one of the subproblems is divided further.
-
-	The next step after branch-and-reduce is bounding, which is able to prune whole sub problems.
-In the following part I will consider the minimization of an objective function without a loss of generality.
-For the bounding it is first necessary to initialize a upper bound with a value, for example with $\infty$.
-This upper bound is updated during the process every time a new optimal solution is found to this new optimum.
-If a new possible solution is found which is above this upper bound the algorithm can prune it, as only the minimum solution is of interest.
-The actually pruning of subproblems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
-If this relaxed optimum for a certain subproblem is greater than the current upper bound it is possible to prune this sub problem, as it can not contain a better solution.
-
-	\begin{figure}
 		\begin{tikzpicture}
@@
 			\draw (f) -- (g) -- (c) -- (b) -- (f);
+			\draw[fill=blue, fill opacity=0.2] (c) -- (d) -- (e) -- (c);
+			\draw[fill=green, fill opacity=0.2] (a) -- (f) -- (g) -- (e) -- (a);
+			\draw[fill=brown, fill opacity=0.2] (f) -- (g) -- (c) -- (b) -- (f);
 			\node at (2.5, 4) {$P_0$};
@@
 		\end{tikzpicture}	
-		\center
-		\caption{Possible candidates for a solution and the related search space}
-		\label{fig:bounding}
+		\caption{Splitting of a search space into three subproblems and the possible candidates for solutions}
+		\label{fig:branchAndBound}
 	\end{figure}
 
-	Consider for this \reffig{fig:bounding}, where already a certain optimum $f_1$ was found in subproblem $P_1$. If now the relaxed optimum $f_{P0}$ is worse than this optimum
+	As the basic concept of the finding solutions is the enumeration of possible candidates it is necessary to seperate the search space in a useful way, which is the task for the branch-and-reduce step.
+Actually this step is very similar to the concept of branching, it also depends on a certain heuristic to divide the search space into two or even more disjoint parts.
+This heuristic can either be selected by the user or be an general one, like splitting the biggest variable domain into two equal sized parts.
+If the basic search space contains a big amount of possible candidate solution one reduction may not be enough and therefore this step is applied recursively to the newly create subproblems.
+This process can be stopped if the number of solutions is reduced to an amount, where it is possible to enumerate all of them and check if they are optimal.
+A possible division into subproblems can be seen in \reffig{fig:branchAndReduce}, where for legibility always only one of the subproblems is divided further.
+
+	The next step after branch-and-reduce is bounding, which is able to prune whole sub problems.
+In the following part I will consider the minimization of an objective function without a loss of generality.
+For the bounding it is first necessary to initialize a upper bound with a value, for example with $\infty$.
+This upper bound is updated during the process every time a new optimal solution is found to this new optimum.
+If a new possible solution is found which is above this upper bound the algorithm can prune it, as only the minimum solution is of interest.
+The actually pruning of subproblems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
+If this relaxed optimum for a certain subproblem is greater than the current upper bound it is possible to prune this sub problem, as it can not contain a better solution.
+
+	Consider for this \reffig{fig:branchAndBound}, where already a certain optimum $f$ was found in subproblem $P_1$.
+If now the relaxed optimum $f_{P0}$ in $P_0$ and $f_{P2}$ in $P_2$ is worse than this optimum
 	\begin{equation}
-		f_{P0} > f_1
+		f_{P0} > f
 	\end{equation}
-	the whole subproblem $P_{0}$ can be pruned and it is not necessary to enumerate the solution candidates in it (or to apply split it further up).
+	\begin{equation}
+		f_{P2} > f
+	\end{equation}
+	the whole subproblems $P_{0}$ and $P_{2}$ can be pruned and it is not necessary to enumerate the solution candidates in them (or to split them further up).
 

### 2014-05-08 paper.tex
reduced two figures to only one
improved figure for branch-and-bound
abstract
conclusion

@@
 	\begin{abstract}	
-		NEXT TRY
+		Constraint programming is a way to describe and solve mathematical problems with arbitrary constraints.
+In the following paper I will explain the implemented algorithms in Gecode, a tool which can be used to solve problems through constraint programming.
 	\end{abstract}
@@
 			\draw[fill=brown, fill opacity=0.2] (f) -- (g) -- (c) -- (b) -- (f);
-			\node at (2.5, 4) {$P_0$};
-			\node at (1, 1) {$P_1$};
-			\node at (3.5, 1.5) {$P_2$};
+			\node at (2.75, 3.75) {$P_0$};
+			\node at (1.25, 1.25) {$P_1$};
+			\node at (3.75, 1.75) {$P_2$};
 		\end{tikzpicture}	
@@
 	\section{Conclusion}
-	Constraint programming is a very useful tool in practice, as it can solve abritrary combinatorial or optimization problems.
-\emph{Gecode}, as representant of the various toolsets, implements a handy interface to transform a mathematical description of a problem into code, which then can be used to solve the problem.
-In conclusion, \emph{Gecode} provides all the necessary tools, combines them with a nice interface, and is still very flexible concering optimzation for a certain area through the possibility to implement custom search strategies, based on some well studied and efficient algorithms.
+	\emph{Gecode}, as representant for an implemention of constraint programming, provides everything necessary to describe and solve arbitrary combinatorial or optimization problems.
+The interface allows an adaption to the specific area through modificiations of algorithms, restarts or the implementation of a custom search algorithm, based on the already provided ones.
+Therefore \emph{Gecode} can be used in various areas, starting from scheduling, over portfolio optimization up to technical problems like placement.
 

### 2014-05-08 paper.tex
improved introduction

@@
 
-\newcommand{\reffig}[1]{{figure \ref{#1}}}
+\newcommand{\reffig}[1]{{figure~\ref{#1}}}
 \newcommand{\refeq}[1]{{(\ref{#1})}}
@@
 	Constraint programming is a method to solve mathematical problems through the formulation of constraints, which describe the desired solution.
-As a formal and exact definition, like it can be found in \cite[p.~16]{handbookCP} for constraint programming, is not as helpful as an example, I would like to give you a short overview on the topic through the problem of the placement of two squares such that they do not overlap.
+A formal and exact definition for constraint programming can be found in \cite[p.~16]{handbookCP}.
+To visualize the topic better I will start with the problem of the placement of two squares such that they do not overlap.
 
@@
 		\end{tikzpicture}
-		\caption{two squares which should not overlap}
+		\caption{Two squares which should not overlap}
 		\label{fig:squares}

### 2014-05-08 paper.tex
corrections

@@
 		Constraint programming is a way to describe and solve mathematical problems with arbitrary constraints.
-In the following paper I will explain the implemented algorithms in Gecode, a tool which can be used to solve problems through constraint programming.
+In the following paper I will explain the implemented algorithms in \emph{Gecode}, a tool which can be used to solve problems through constraint programming.
 	\end{abstract}
@@
 	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:squares} in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected algorithm.
-If there would be only linear constraints like it would be possible to solve this problem through linear programming \cite[1]{linearProgramming}.
+If there would be only linear constraints like \refeq{eq:linear} it would be even possible to solve the problem through linear programming \cite[1]{linearProgramming}.
 As linear programming is only a subset of constraint programming the latter one is the more general one with more practical applications; not every problem in the real world is linear or can be linearized in a useful way.
-Consequently, constraint programming with its ability to consider constraints in their most general form can be used in various fields like scheduling of resources, computer-aided design and robotics \cite[p.~221]{trendsInCP}.
+Consequently, constraint programming with its ability to consider constraints in their most general form can be used in various fields like scheduling of resources, computer-aided design or robotics \cite[p.~221]{trendsInCP}.
 
@@
 		c_2: y - x - 1 = 0
+		\label{eq:linear}
 	\end{equation}

### 2014-05-08 paper.tex
corrections

@@
 	As constraint programming is the generalization of linear programming it is at least possible to define linear constraints.
-Which statements are possible to make and how handy it is to do so varies from the programming language in which the problem is implemented and the used toolset.
+Which statements are possible to make and how comfortable it is to do so varies from the programming language in which the problem is implemented and the used toolset.
 Typical constraints are
@@
 	Different constraint types can also be combined, for example inequalities or equalities through logic operators to boolean expressions.
-Some of these constraints may acquire special attention during the search for solutions to improve the execution reason.
+Some of these constraints may acquire special attention during the search for solutions to improve the runtime of the search.
 For this reason a lot of work has been done for example on the alldifferent constraint \cite{allDifferent}, which would be such a combined constraint.
@@
 
-	In constraint programming search trees, like the one in \reffig{fig:searchTree}, describe the partition of the search space and their nodes have names, depending on position in the tree.
+	In constraint programming search trees, like the one in \reffig{fig:searchTree}, describe the partition of the search space and their nodes have names, depending on their position in the tree.
 The root node is the one from where the search is started and all variable domains are still complete.
-This one is also the first one with a branch, where for the first time the search space is split up into disjoint sets.
+This one is also the first one with a branch, where the search space is split up into disjoint sets.
 This approach is similiar to the divide-and-conquer technique as it reduces a complex problem into two or more parts with less complexity.
@@
 	Without constraint propagation and branching first on $x$ the search trees turns out to be the one in \reffig{fig:bigTree}.
-Obviously the right parts of this tree can already be pruned already at the root by propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these subtrees like in \reffig{fig:smallTree}.
+Obviously the right parts of this tree can be already pruned at the root by propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these subtrees.
+With constraint propagation fewer branches have to be made, like it can be seen in \reffig{fig:smallTree}.
 
-	The actual implementation of constraint propagation differs again from one toolset to another, but theory underneath is the same.
+	The actual implementation of constraint propagation differs again from one toolset to another, but the theory underneath is the same.
 As the topic of this paper is the implementation in \emph{Gecode} I will concentrate on this one.
@@
 	Constraints in \emph{Gecode} are always evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the subtrees.
-The problem here is to decide in which order the constraints should be considered, and it may even be necessary to evaluate some constraints more often.
-A short example will explain this very well \cite[p.~575]{handbookCP}: Let there be two constraints
+The problem here is to decide in which order the constraints should be considered, and it may be even necessary to evaluate some constraints more often.
+A short example will explain this better \cite[p.~575]{handbookCP}: Let there be two constraints
 	\begin{equation}
@@
 	\end{equation}
-	and let the variable domains be $x, y \in [-4, 4]$. Starting with the evaluation of constraint one follows a chain of reductions of the variable domains, based only on constraint propagation:
+	and let the variable domains be $x, y \in [-4, 4]$. Starting with the evaluation of constraint $c_1$ follows a chain of reductions of the variable domains, based only on constraint propagation:
 	\begin{equation}
@@
 	The next decision to be made is which constraint should be evaluated first.
-In the example above the constraints were quite simple and therefore it won't make a big difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time.
-Unfortunately not all constraints are that easy to evaluate and for this case \emph{Gecode} defines for every constraint a certain cost \cite[p.~275]{programmingGecode}.
+In the example above the constraints were quite simple and therefore it would not make a big difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time.
+Unfortunately not all constraints are that fast to evaluate and for this case \emph{Gecode} defines for every constraint a certain cost \cite[p.~275]{programmingGecode}.
 Based on this cost the constraints are sorted into buckets which are then iterated from the constraints with lower costs to the ones with higher costs.

### 2014-05-08 paper.tex
corrections

@@
 	\subsection{Search Algorithms}
-	In constraint programming search algorithms define the way the search tree is constructed, starting from the root node.
+	In constraint programming search algorithms define the way the search tree is constructed and traversed, starting from the root node.
 To improve the performance of the search usually at every newly created node a constraint propagation is executed to narrow the search space down.
 Based on these two steps then either one, the best or no solution is found.
-What the actual result in the end is depends on the type of search algorithm.
+What the actual result in the end is depends on the type of the search algorithm.
 
-	In \emph{Gecode} it is possible to define custom strategies, but the user is supported at this point with some standard engines which can be used to build custom ones:
+	In \emph{Gecode} it is possible to define custom strategies, but the user is supported at this point with some standard algorithms which can be used to build custom ones:
 	\begin{itemize}
-	\item Depth-First Search
-	\item Limited Discrepancy Search
-	\item Branch-and-Bound Search
-	\item Depth-First Search Restart Optimization
+		\item Depth-first search
+		\item Limited Discrepancy search
+		\item Branch-and-bound search
+		\item Depth-first search restart optimization
 	\end{itemize}
 
-	In the following part I will discuss branching strategies, the Depth-first search and one of its variants, the Branch-and-bound search, the Limited-Discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
+	In the following part I will discuss branching strategies, the Depth-first search, the Branch-and-bound search, the Limited-Discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
 
 	\subsubsection{Branching Strategies}
-	If at one node it is not possible to empty one variable domain or to reduce all variable domains to the size one it is necessary to make a decision: How should the next subtree be constructed?
+	If at one node it is not possible to empty one variable domain or to reduce all variable domains to the size of one it is necessary to make a decision: How should the next subtree be constructed?
 The question is answered by the selected branching strategy.
 No matter which one is selected, they all have in common that they create additional, temporary constraints, which are only applied to the specific subtrees.
-This step is called \emph{posting} a constraint on a branch.
+This process is called \emph{posting} a constraint on a branch.
 
@@
 		\item Binary choice points: One value $x_i$ of the domain $D$ of the variable $x$ is selected and the domain is split up into two subtrees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted.
-This results into something like a binary tree at this point.
+This results into something similar to a binary tree at this point.
 		\item Domain splitting: The domain $D$ for the variable $x$ is split up into two disjoint intervals or sets, for example $x < 2$ and $x \ge 2$.
@@
 
-	As for variables with a Float domain the Enumeration and Binary choice points are not feasible, the last one, domain splitting, is typically used.
+	As for variables with a Float domain the enumeration and binary choice points are not feasible, the last one, domain splitting, is typically used.
 As termination criteria for a Float it is often necessary to define a certain epsilon $\epsilon \in \mathbb{R}_{>0}$ which describes the interval length, at which the solver states to have found a solution.
 
-	For all branching strategies there is a choice of freedom left, which can be used to implement some heuristics.
-This additional information is either gained during the search itself of predefined by the user.
+	For branching strategies there is a choice of freedom left, for instance where the interval is split up, and this freedom can be used to implement some heuristics.
+This additional information is either gained during the search itself or predefined by the user.
 
 	\subsubsection{Depth-First Search}
-	The Depth-first search is a good point to start, as it was also historical the first developed one.
+	The Depth-first search was historical the first one which was developed.
 The idea behind this search algorithm is that every possible combination of variables is examined.
@@
 	The starting point is the root node, from where on the algorithm is called recursively.
-At every node mainly two steps are executed: First a constraint propagation and then a branching, which divides the search space into two disjoint sub spaces, on which the same algorithm is applied again.
+At every node mainly two steps are executed: First a constraint propagation to reduce the variable domains and second a branching, which divides the search space into two disjoint sub spaces, on which the same algorithm is applied again.
 The search stops in a leave if either no branching is possible anymore and therefore a valid solution is found, or one or more variable domains are emptied, which indicates that in this leave no possible solution exists.
@@
 	\subsubsection{Depth-First Search with Optimization}
-	If in the constraint programming an objective is defined which should be either minimized or maximized a modified version of the Depth-first search can be applied.
+	If in the constraint programming an objective is defined which should be either minimized or maximized, a modified version of the Depth-first search can be applied.
 The only difference is that it evalutes the objective function for every found solution concerning the constraints and stores during the search only the best one.
@@
 	A branch-and-bound search is an algorithm for optimization of an objective function under certain constraints and only discrete possible candidates for a solution.
-To fulfill the constraints the first step, branch-and-reduce is applied, and the second one, bounding, is used to prune whole areas with non-optimal solutions.
+To fulfill the constraints the first step, branch-and-reduce is, applied, and the second one, bounding, is used to prune whole areas with non-optimal solutions.
 
@@
 
-	As the basic concept of the finding solutions is the enumeration of possible candidates it is necessary to seperate the search space in a useful way, which is the task for the branch-and-reduce step.
+	As the basic concept for finding solutions in Branch-and-bound is an enumeration of all possible candidates, it is necessary to seperate the search space in a useful way, which is the task for the Branch-and-reduce step.
 Actually this step is very similar to the concept of branching, it also depends on a certain heuristic to divide the search space into two or even more disjoint parts.
-This heuristic can either be selected by the user or be an general one, like splitting the biggest variable domain into two equal sized parts.
-If the basic search space contains a big amount of possible candidate solution one reduction may not be enough and therefore this step is applied recursively to the newly create subproblems.
+This heuristic can either be selected by the user or be a general one, like splitting the biggest variable domain into two equal sized parts.
+If the basic search space contains a big amount of possible candidate solutions one reduction may not be enough and therefore this step is applied recursively to the newly create subproblems.
 This process can be stopped if the number of solutions is reduced to an amount, where it is possible to enumerate all of them and check if they are optimal.
-A possible division into subproblems can be seen in \reffig{fig:branchAndReduce}, where for legibility always only one of the subproblems is divided further.
+A possible division into subproblems can be seen in \reffig{fig:branchAndBound}, where for legibility Branch-and-reduce was applied only twice: The first execution split the problem into $P_0$ and the second split the other half into $P_1$ and $P_2$.
 
-	The next step after branch-and-reduce is bounding, which is able to prune whole sub problems.
+	The next step after Branch-and-reduce is the Bounding, which is able to prune whole sub problems.
 In the following part I will consider the minimization of an objective function without a loss of generality.
-For the bounding it is first necessary to initialize a upper bound with a value, for example with $\infty$.
+For the Bounding it is first necessary to initialize a upper bound with a value, for example with $\infty$.
 This upper bound is updated during the process every time a new optimal solution is found to this new optimum.
@@
 The actually pruning of subproblems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
-If this relaxed optimum for a certain subproblem is greater than the current upper bound it is possible to prune this sub problem, as it can not contain a better solution.
+If this relaxed optimum for a certain subproblem is greater than the current upper bound it is possible to prune this complete sub problem, as it can not contain a better solution.
 
@@
 	\subsubsection{Limited Discrepancy Search}
-	Limited Discrepancy Search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy}.
+	Limited Discrepancy search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy}.
 It depends on some heurisitics for the tree, which are determined through branching strategies.
 The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive.
-Second, heuristics work worse the farther away from the solution they are applied, therefore decisions close to the root node are possible to fail.
+Second, heuristics work worse the farther away from the solution they are applied, therefore decisions close to the root node are likely to fail.
 Consequently, in a chronological backtrack search the worst decisions are the most expensive ones.
-The Limited Discrepancy Search solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later.
-The improvement compared to a chronological backtrack was shown by experimental results and theoretical proven in \cite{limitedDiscrepancy}.
+The Limited Discrepancy search solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later.
+The improvement compared to a chronological backtrack was shown by experimental results and theoretically proven in \cite{limitedDiscrepancy}.
 
 	\subsubsection{Backjumping}
-	A specific version of backjumping was already mentioned, the Limited Discrepancy Search, where instead of jumping one node back the algorithm always jumps as far back to the root node as possible.
-In general this is proven to be better than the chronological backtracking \cite{limitedDiscrepancy}, but for certain problems it may be possible to find better points to jump back.
+	A specific version of backjumping was already mentioned, the Limited Discrepancy search, where instead of jumping one node back the algorithm always jumps as far back to the root node as possible.
+In general this is proven to be better than the chronological backtracking \cite{limitedDiscrepancy}, but for certain problems it may be possible to find better nodes to jump back.
 A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the deadend.
@@
 	\subsubsection{Restart Strategies}
-	Branching means always to apply certain heuristics and these heuristics may fail or produce bad results for the first execution.
+	Branching always means to apply certain heuristics and these heuristics may fail or produce bad results for the first execution of the search algorithm.
 If it turns out that the search tree is build in a disadvantageous way a restart strategy may do the trick to improve the search.

### 2014-05-08 paper.tex
corrections

@@
 	\markboth{Advanced Seminar for Security in Information Technology, Summer Term 2014}%
-	{Benedikt Schmidt: Constraint Programming - Inside Gecode}	
+	{Benedikt Schmidt: Constraint Programming - Inside \emph{Gecode}}	
 	\maketitle
@@
 
-	In the following part I will discuss branching strategies, the Depth-first search, the Branch-and-bound search, the Limited-Discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
+	In the following part I will discuss branching strategies, the Depth-first search, the Branch-and-bound search, the Limited Discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
 
@@
 	\subsubsection{Limited Discrepancy Search}
-	Limited Discrepancy search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy}.
+	The Limited Discrepancy search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy}.
 It depends on some heurisitics for the tree, which are determined through branching strategies.

### 2014-05-08 paper.tex
moved the branching strategies out of the part about search algorithms

@@
 
-	\subsection{Search Algorithms}
-	In constraint programming search algorithms define the way the search tree is constructed and traversed, starting from the root node.
-To improve the performance of the search usually at every newly created node a constraint propagation is executed to narrow the search space down.
-Based on these two steps then either one, the best or no solution is found.
-What the actual result in the end is depends on the type of the search algorithm.
-
-	In \emph{Gecode} it is possible to define custom strategies, but the user is supported at this point with some standard algorithms which can be used to build custom ones:
-	\begin{itemize}
-		\item Depth-first search
-		\item Limited Discrepancy search
-		\item Branch-and-bound search
-		\item Depth-first search restart optimization
-	\end{itemize}
-
-	In the following part I will discuss branching strategies, the Depth-first search, the Branch-and-bound search, the Limited Discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
-
-	\subsubsection{Branching Strategies}
+		\subsection{Branching Strategies}
 	If at one node it is not possible to empty one variable domain or to reduce all variable domains to the size of one it is necessary to make a decision: How should the next subtree be constructed?
@@
 This additional information is either gained during the search itself or predefined by the user.
+
+	\subsection{Search Algorithms}
+	In constraint programming search algorithms define the way the search tree is constructed and traversed, starting from the root node.
+To improve the performance of the search usually at every newly created node a constraint propagation is executed to narrow the search space down.
+Based on these two steps then either one, the best or no solution is found.
+What the actual result in the end is depends on the type of the search algorithm.
+
+	In \emph{Gecode} it is possible to define custom strategies, but the user is supported at this point with some standard algorithms which can be used to build custom ones:
+	\begin{itemize}
+		\item Depth-first search
+		\item Limited Discrepancy search
+		\item Branch-and-bound search
+		\item Depth-first search restart optimization
+	\end{itemize}
+
+	In the following part I will discuss the Depth-first search, the Branch-and-bound search, the Limited Discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
 

### 2014-05-08 paper.tex
extended introduction

@@
 
-	In the following I will describe the necessary steps to solve a problem with constraint programming and as example of an implementation I will always refer to \emph{Gecode} \cite{gecode}.
+	In the following I will describe the necessary steps to describe and solve a problem with constraint programming: Variables, constraints, search trees, constraint propagation and search algorithms.
+As example of an implementation I will always refer to \emph{Gecode} \cite{gecode}.
 

### 2014-05-08 paper.tex
spell check

@@
 
-	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:squares} in their most natural form as equations and inequations and will find all possible solutions or just one of them, depending on the selected algorithm.
+	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:squares} in their most natural form as equations and in-equations and will find all possible solutions or just one of them, depending on the selected algorithm.
 If there would be only linear constraints like \refeq{eq:linear} it would be even possible to solve the problem through linear programming \cite[1]{linearProgramming}.
@@
 	In constraint programming a variable is connected to a set of possible values, the so called domain.
-Depending on the domain a variable is called an integer, boolean or floating variable.
+Depending on the domain a variable is called an integer, Boolean or floating variable.
 Possible ranges for a variable $x$ of each domain are:
 	\begin{itemize}
-		\item Integer: finite or inifnite sets like $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$, \dots
+		\item Integer: finite or infinite sets like $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$, \dots
 		\item Bool: $x \in \{\text{true}, \text{false}\}$
@@
 	As constraint programming is the generalization of linear programming it is at least possible to define linear constraints.
-Which statements are possible to make and how comfortable it is to do so varies from the programming language in which the problem is implemented and the used toolset.
+Which statements are possible to make and how comfortable it is to do so varies from the programming language in which the problem is implemented and the used tool set.
 Typical constraints are
@@
 
-	Different constraint types can also be combined, for example inequalities or equalities through logic operators to boolean expressions.
+	Different constraint types can also be combined, for example inequalities or equalities through logic operators to Boolean expressions.
 Some of these constraints may acquire special attention during the search for solutions to improve the runtime of the search.
@@
 This one is also the first one with a branch, where the search space is split up into disjoint sets.
-This approach is similiar to the divide-and-conquer technique as it reduces a complex problem into two or more parts with less complexity.
+This approach is similar to the divide-and-conquer technique as it reduces a complex problem into two or more parts with less complexity.
 The splitting, which adds additional constraints on the variables, is then recursively done with different heuristics to partition the search space into smaller parts.
-If one of the parts is small enough to decide if it is feasible the node can be called a either a deadend or a solution.
+If one of the parts is small enough to decide if it is feasible the node can be called a either a dead end or a solution.
 To end up as the dead end there must be no feasible solutions left in the search space at this node.
@@
 The term propagation refers in this area to the reduction of domains of variables which are infeasible, considering other constraints and affected variable domains.
-As the size of the variable domain is related to possibilities which have to be checked by a search, the constraint propagation is able to prune complete infeasible subtrees.
+As the size of the variable domain is related to possibilities which have to be checked by a search, the constraint propagation is able to prune complete infeasible sub trees.
 Therefore the whole search process can be accelerated through constraint propagation.
@@
 	Without constraint propagation and branching first on $x$ the search trees turns out to be the one in \reffig{fig:bigTree}.
-Obviously the right parts of this tree can be already pruned at the root by propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these subtrees.
+Obviously the right parts of this tree can be already pruned at the root by propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these sub trees.
 With constraint propagation fewer branches have to be made, like it can be seen in \reffig{fig:smallTree}.
 
-	The actual implementation of constraint propagation differs again from one toolset to another, but the theory underneath is the same.
+	The actual implementation of constraint propagation differs again from one tool set to another, but the theory underneath is the same.
 As the topic of this paper is the implementation in \emph{Gecode} I will concentrate on this one.
 
-	Constraints in \emph{Gecode} are always evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the subtrees.
+	Constraints in \emph{Gecode} are always evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the sub trees.
 The problem here is to decide in which order the constraints should be considered, and it may be even necessary to evaluate some constraints more often.
@@
 The question is answered by the selected branching strategy.
-No matter which one is selected, they all have in common that they create additional, temporary constraints, which are only applied to the specific subtrees.
+No matter which one is selected, they all have in common that they create additional, temporary constraints, which are only applied to the specific sub trees.
 This process is called \emph{posting} a constraint on a branch.
 
-	The branching strategy depends on the variable domain: For Integer (and Boolean) variables there are three common strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the subtrees.
+	The branching strategy depends on the variable domain: For Integer (and Boolean) variables there are three common strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the sub trees.
 	\begin{itemize}
 		\item Enumeration: The domain for $x \in D = {x_1, x_2, \dots}$ is split up into $|D|$-branches where for every branch $i$ one additional constraint $c:\ x = x_i$ is posted.
-		\item Binary choice points: One value $x_i$ of the domain $D$ of the variable $x$ is selected and the domain is split up into two subtrees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted.
+		\item Binary choice points: One value $x_i$ of the domain $D$ of the variable $x$ is selected and the domain is split up into two sub trees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted.
 This results into something similar to a binary tree at this point.
@@
 The idea behind this search algorithm is that every possible combination of variables is examined.
-Obviously in its basic form it can not be applied to problems with inifinite domains; for problems with such variable domains it is necessary to use for instance Branch-and-reduce.
+Obviously in its basic form it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance Branch-and-reduce.
 Sometimes the Depth-first search is also referred to as chronological backtrack search.
@@
 	This search is a so-called complete algorithm as it can be proven that all possible solutions will be found \cite[p.~85]{handbookCP}.
-Because of this characteristic it is very useful to proove for example that for a certain problem no solution exists.
+Because of this characteristic it is very useful to prove for example that for a certain problem no solution exists.
 The main drawback obviously is the execution time, as it does an exhaustive search, and maybe even the memory usage if during the search many solutions are found.
@@
 	If in the constraint programming an objective is defined which should be either minimized or maximized, a modified version of the Depth-first search can be applied.
-The only difference is that it evalutes the objective function for every found solution concerning the constraints and stores during the search only the best one.
+The only difference is that it evaluates the objective function for every found solution concerning the constraints and stores during the search only the best one.
 To accelerate the search it is possible to add an additional constraint which defines that the new objective must better than the best so far.
@@
 		\end{tikzpicture}	
-		\caption{Splitting of a search space into three subproblems and the possible candidates for solutions}
+		\caption{Splitting of a search space into three sub problems and the possible candidates for solutions}
 		\label{fig:branchAndBound}
@@
 
-	As the basic concept for finding solutions in Branch-and-bound is an enumeration of all possible candidates, it is necessary to seperate the search space in a useful way, which is the task for the Branch-and-reduce step.
+	As the basic concept for finding solutions in Branch-and-bound is an enumeration of all possible candidates, it is necessary to separate the search space in a useful way, which is the task for the Branch-and-reduce step.
 Actually this step is very similar to the concept of branching, it also depends on a certain heuristic to divide the search space into two or even more disjoint parts.
 This heuristic can either be selected by the user or be a general one, like splitting the biggest variable domain into two equal sized parts.
-If the basic search space contains a big amount of possible candidate solutions one reduction may not be enough and therefore this step is applied recursively to the newly create subproblems.
+If the basic search space contains a big amount of possible candidate solutions one reduction may not be enough and therefore this step is applied recursively to the newly create sub problems.
 This process can be stopped if the number of solutions is reduced to an amount, where it is possible to enumerate all of them and check if they are optimal.
-A possible division into subproblems can be seen in \reffig{fig:branchAndBound}, where for legibility Branch-and-reduce was applied only twice: The first execution split the problem into $P_0$ and the second split the other half into $P_1$ and $P_2$.
+A possible division into sub problems can be seen in \reffig{fig:branchAndBound}, where for legibility Branch-and-reduce was applied only twice: The first execution split the problem into $P_0$ and the second split the other half into $P_1$ and $P_2$.
 
@@
 If a new possible solution is found which is above this upper bound the algorithm can prune it, as only the minimum solution is of interest.
-The actually pruning of subproblems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
-If this relaxed optimum for a certain subproblem is greater than the current upper bound it is possible to prune this complete sub problem, as it can not contain a better solution.
+The actually pruning of sub problems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
+If this relaxed optimum for a certain sub problem is greater than the current upper bound it is possible to prune this complete sub problem, as it can not contain a better solution.
 
-	Consider for this \reffig{fig:branchAndBound}, where already a certain optimum $f$ was found in subproblem $P_1$.
+	Consider for this \reffig{fig:branchAndBound}, where already a certain optimum $f$ was found in sub problem $P_1$.
 If now the relaxed optimum $f_{P0}$ in $P_0$ and $f_{P2}$ in $P_2$ is worse than this optimum
@@
 	\end{equation}
-	the whole subproblems $P_{0}$ and $P_{2}$ can be pruned and it is not necessary to enumerate the solution candidates in them (or to split them further up).
+	the whole sub problems $P_{0}$ and $P_{2}$ can be pruned and it is not necessary to enumerate the solution candidates in them (or to split them further up).
 
-	Combined with appropriate branching the bounding can prune several candiates at once, without the necessity to enumerate them. Therefore bounding is able to improve the performance of the search.
+	Combined with appropriate branching the bounding can prune several candidates at once, without the necessity to enumerate them. Therefore bounding is able to improve the performance of the search.
 
@@
 	The Limited Discrepancy search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy}.
-It depends on some heurisitics for the tree, which are determined through branching strategies.
+It depends on some heuristics for the tree, which are determined through branching strategies.
 The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive.
@@
 In general this is proven to be better than the chronological backtracking \cite{limitedDiscrepancy}, but for certain problems it may be possible to find better nodes to jump back.
-A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the deadend.
+A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the dead end.
 How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is done wrong the backjumping can even be less efficient than chronological backtracking \cite[p.~100]{handbookCP}.
@@
 	\section{Conclusion}
-	\emph{Gecode}, as representant for an implemention of constraint programming, provides everything necessary to describe and solve arbitrary combinatorial or optimization problems.
-The interface allows an adaption to the specific area through modificiations of algorithms, restarts or the implementation of a custom search algorithm, based on the already provided ones.
+	\emph{Gecode}, as representative for an implementation of constraint programming, provides everything necessary to describe and solve arbitrary combinatorial or optimization problems.
+The interface allows an adaption to the specific area through modifications of algorithms, restarts or the implementation of a custom search algorithm, based on the already provided ones.
 Therefore \emph{Gecode} can be used in various areas, starting from scheduling, over portfolio optimization up to technical problems like placement.

### 2014-05-09 paper.tex
corrections Michael

@@
 
-\newcommand{\reffig}[1]{{figure~\ref{#1}}}
+\newcommand{\reffig}[1]{{Fig.~\ref{#1}}}
 \newcommand{\refeq}[1]{{(\ref{#1})}}
@@
 	\begin{abstract}	
-		Constraint programming is a way to describe and solve mathematical problems with arbitrary constraints.
-In the following paper I will explain the implemented algorithms in \emph{Gecode}, a tool which can be used to solve problems through constraint programming.
+		Constraint programming is a method to describe and solve mathematical problems with arbitrary constraints.
+In the following paper, I will explain the implemented algorithms in \emph{Gecode}, a tool which can be used to solve problems through constraint programming.
 	\end{abstract}
@@
 A formal and exact definition for constraint programming can be found in \cite[p.~16]{handbookCP}.
-To visualize the topic better I will start with the problem of the placement of two squares such that they do not overlap.
+For a better visualization of the topic, I will start with the problem of the placement of two squares such that they do not overlap.
 
@@
 	\end{equation}
-	The task for the constraint programming tool at this point is then to find feasible solutions which fulfill the constraints.
+	The task for the constraint programming tool is to find feasible solutions which fulfill the constraints.
 
-	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:squares} in their most natural form as equations and in-equations and will find all possible solutions or just one of them, depending on the selected algorithm.
-If there would be only linear constraints like \refeq{eq:linear} it would be even possible to solve the problem through linear programming \cite[1]{linearProgramming}.
-As linear programming is only a subset of constraint programming the latter one is the more general one with more practical applications; not every problem in the real world is linear or can be linearized in a useful way.
+	A solver for constraint programming like \emph{Gecode} provides the necessary tools to define constraints like \refeq{eq:squares} in their most natural form as equalities and inequalities.
+The solver will find all possible solutions or just one of them, depending on the selected algorithm.
 Consequently, constraint programming with its ability to consider constraints in their most general form can be used in various fields like scheduling of resources, computer-aided design or robotics \cite[p.~221]{trendsInCP}.
 
-	In the following I will describe the necessary steps to describe and solve a problem with constraint programming: Variables, constraints, search trees, constraint propagation and search algorithms.
-As example of an implementation I will always refer to \emph{Gecode} \cite{gecode}.
+	In the following I will explain the necessary steps to describe and solve a problem with constraint programming: Variables, constraints, search trees, constraint propagation and search algorithms.
+As example of an implementation I will refer to \emph{Gecode} \cite{gecode}.
 
@@
 	\subsection{Variable Domains}
-	In constraint programming a variable is connected to a set of possible values, the so called domain.
-Depending on the domain a variable is called an integer, Boolean or floating variable.
+	In constraint programming, a variable is connected to a set of possible values, the so called domain.
+Depending on the domain, a variable is called an integer, Boolean or floating variable.
 Possible ranges for a variable $x$ of each domain are:
@@
 	\end{itemize}
-	This distinction is fundamental as especially for floating variables the classical combinatorial approach to constraint programming is impossible.
+	This distinction is fundamental as especially for floating variables a combinatorial approach to constraint programming is impossible.
 
 	\subsection{Constraint Types}
-
-	As constraint programming is the generalization of linear programming it is at least possible to define linear constraints.
-Which statements are possible to make and how comfortable it is to do so varies from the programming language in which the problem is implemented and the used tool set.
+	Which statements are possible to make and how comfortable it is to do so varies from the programming language in which the problem is implemented.
+For sure, there is also a difference in the interfaces of the libraries for constraint programming.
 Typical constraints are
@@
 Some of these constraints may acquire special attention during the search for solutions to improve the runtime of the search.
-For this reason a lot of work has been done for example on the alldifferent constraint \cite{allDifferent}, which would be such a combined constraint.
+For this reason a lot of work has been done for example on the alldifferent constraint \refeq{eq:alldifferent} \cite{allDifferent}, which would be such a combined constraint.
 	\begin{equation}
@@
 
-	In constraint programming search trees, like the one in \reffig{fig:searchTree}, describe the partition of the search space and their nodes have names, depending on their position in the tree.
-The root node is the one from where the search is started and all variable domains are still complete.
-This one is also the first one with a branch, where the search space is split up into disjoint sets.
+	In constraint programming, search trees, like the one in \reffig{fig:searchTree}, describe the partition of the search space.
+Their nodes have names, depending on their position in the tree.
+The search is started from the root node.
+At this point, all variable domains are still complete.
+The root node is also the first one with a branch, where the search space is split up into disjoint sets.
 This approach is similar to the divide-and-conquer technique as it reduces a complex problem into two or more parts with less complexity.
@@
 If one of the parts is small enough to decide if it is feasible the node can be called a either a dead end or a solution.
-To end up as the dead end there must be no feasible solutions left in the search space at this node.
+To end up as a dead end, there must be no feasible solutions left in the search space at this node.
 To turn the node into a solution it is necessary that all members of the reduced search space are valid considering the constraints.
+This can be the case if either only one choice is left for every variable, and the constraints are valid for this choice, or if a whole interval of a floating variable is valid for the constraints.
 
@@
 	Constraint propagation is a key concept in constraint programming as it reduces the size of the search space significantly.
-The term propagation refers in this area to the reduction of domains of variables which are infeasible, considering other constraints and affected variable domains.
-As the size of the variable domain is related to possibilities which have to be checked by a search, the constraint propagation is able to prune complete infeasible sub trees.
-Therefore the whole search process can be accelerated through constraint propagation.
+The term propagation refers in this area to the reduction of choices for variables which are infeasible, considering other constraints and affected variable domains.
+As the size of the variable domain is related to the possibilities which have to be checked by a search through additional subtrees, the constraint propagation is able to prune complete infeasible sub trees.
+Therefore, the whole search process can be accelerated through constraint propagation.
 
@@
 
-	To visualize this effect consider a constraint problem based on two variables $x \in \{1, 5, 6\}$ and $y \in \{1, 2\}$ and one constraint based on them.
+	To illustrate this effect consider a constraint problem based on two variables $x \in \{1, 5, 6\}$ and $y \in \{1, 2\}$ and one constraint based on them.
 	\begin{equation}
@@
 	\end{equation}
-	Without constraint propagation and branching first on $x$ the search trees turns out to be the one in \reffig{fig:bigTree}.
-Obviously the right parts of this tree can be already pruned at the root by propagation of constraint $c_1$ and therefore it is not necessary to create and evaluate these sub trees.
+	Without constraint propagation and branching first on $x$, a decision, which may have been made by a certain branching strategy, the search trees turns out to be the one in \reffig{fig:bigTree}.
+By inspection one sees that the right parts of this tree can be already pruned at the root by propagation of constraint $c_1$.
+Therefore, it is not necessary to create and evaluate these sub trees.
 With constraint propagation fewer branches have to be made, like it can be seen in \reffig{fig:smallTree}.
 
-	The actual implementation of constraint propagation differs again from one tool set to another, but the theory underneath is the same.
-As the topic of this paper is the implementation in \emph{Gecode} I will concentrate on this one.
+	The implementation of constraint propagation is different in each tool. However, the same principles are used. In this paper, I will describe the \emph{Gecode} implementation.
 
-	Constraints in \emph{Gecode} are always evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the sub trees.
+	Constraints in \emph{Gecode} are evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the sub trees.
 The problem here is to decide in which order the constraints should be considered, and it may be even necessary to evaluate some constraints more often.
-A short example will explain this better \cite[p.~575]{handbookCP}: Let there be two constraints
+A short example will illustrate this \cite[p.~575]{handbookCP}: Let there be two constraints
 	\begin{equation}
@@
 	To implement this repeated evaluation of constraints in \emph{Gecode} all constraints subscribe to the domains of the variables which are part of them.
-With this method the constraints get noticed whenever a variable domain is reduced and they can signal the system that they need to be evaluated once again.
+With this method, the constraints get notified whenever a variable domain is reduced and they can signal the system that they need to be evaluated once again.
 
@@
 In the example above the constraints were quite simple and therefore it would not make a big difference to evaluate first $c_1$ or $c_2$, the search will take nearly the same time.
-Unfortunately not all constraints are that fast to evaluate and for this case \emph{Gecode} defines for every constraint a certain cost \cite[p.~275]{programmingGecode}.
-Based on this cost the constraints are sorted into buckets which are then iterated from the constraints with lower costs to the ones with higher costs.
+Unfortunately, not all constraints are equally fast to evaluate and for this case \emph{Gecode} defines for every constraint a certain cost \cite[p.~275]{programmingGecode}.
+Based on this cost the constraints are sorted into buckets which are then evaluated in order of ascending cost.
 This means, whenever a variable domain is changed, the first try is to use constraints with the lowest possible cost and only if none of them needs to be evaluated constraints with higher costs are considered.
@@
 
-	The branching strategy depends on the variable domain: For Integer (and Boolean) variables there are three common strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the sub trees.
+	The branching strategy depends on the variable domain: For integer variables there are three common strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the sub trees.
 	\begin{itemize}
@@
 
-	As for variables with a Float domain the enumeration and binary choice points are not feasible, the last one, domain splitting, is typically used.
-As termination criteria for a Float it is often necessary to define a certain epsilon $\epsilon \in \mathbb{R}_{>0}$ which describes the interval length, at which the solver states to have found a solution.
+	As for variables with a float domain the enumeration and binary choice points are not feasible, the last one, domain splitting, is typically used.
+As termination criteria for a float it is often necessary to define a certain epsilon $\epsilon > 0$ which describes the interval length, at which the solver states to have found a solution.
 
@@
 	\subsection{Search Algorithms}
-	In constraint programming search algorithms define the way the search tree is constructed and traversed, starting from the root node.
-To improve the performance of the search usually at every newly created node a constraint propagation is executed to narrow the search space down.
-Based on these two steps then either one, the best or no solution is found.
-What the actual result in the end is depends on the type of the search algorithm.
+	Search algorithms define the way the search tree is constructed and traversed, starting from the root node.
+To improve the performance of the search at every node a constraint propagation is executed, which may narrow the search space down.
+Then, based on these two steps, either one, the best or no solution is found.
+The actual result of the search depends on the type of the search algorithm.
 
-	In \emph{Gecode} it is possible to define custom strategies, but the user is supported at this point with some standard algorithms which can be used to build custom ones:
+	In \emph{Gecode}, it is possible to define custom strategies. However, the user is supported at this point with some standard algorithms which he can use to build custom ones:
 	\begin{itemize}
 		\item Depth-first search
-		\item Limited Discrepancy search
+		\item Limited discrepancy search
 		\item Branch-and-bound search
@@
 
-	In the following part I will discuss the Depth-first search, the Branch-and-bound search, the Limited Discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
+	In the following part, I will discuss the depth-first search, the branch-and-bound search, the limited discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
 
 	\subsubsection{Depth-First Search}
-	The Depth-first search was historical the first one which was developed.
+	The depth-first search was historical the first one which was developed.
 The idea behind this search algorithm is that every possible combination of variables is examined.
-Obviously in its basic form it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance Branch-and-reduce.
-Sometimes the Depth-first search is also referred to as chronological backtrack search.
+In its basic form, it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance branch-and-reduce.
+Sometimes the depth-first search is also referred to as chronological backtrack search.
 
-	The starting point is the root node, from where on the algorithm is called recursively.
-At every node mainly two steps are executed: First a constraint propagation to reduce the variable domains and second a branching, which divides the search space into two disjoint sub spaces, on which the same algorithm is applied again.
-The search stops in a leave if either no branching is possible anymore and therefore a valid solution is found, or one or more variable domains are emptied, which indicates that in this leave no possible solution exists.
+	The starting point for the depth-first search is the root node, from where on the algorithm is called recursively.
+At every node, mainly two steps are executed: First, a constraint propagation to reduce the variable domains and second a branching, which divides the search space into two disjoint sub spaces, on which the same algorithm is applied again.
+The search stops in a leaf of the search tree if either no branching is possible and therefore a valid solution is found, or one or more variable domains are emptied, which indicates that in this leave no possible solution exists.
 
 	This search is a so-called complete algorithm as it can be proven that all possible solutions will be found \cite[p.~85]{handbookCP}.
-Because of this characteristic it is very useful to prove for example that for a certain problem no solution exists.
-The main drawback obviously is the execution time, as it does an exhaustive search, and maybe even the memory usage if during the search many solutions are found.
+Because of this characteristic, it is very useful to prove for example that for a certain problem no solution exists.
+The main drawback is the execution time, as it does an exhaustive search.
+Maybe even the memory usage can become a problem if during the search many solutions are found.
 
-	By stopping the search after one feasible solution is found the algorithm can often be accelerated.
+	By stopping the search after the first feasible solution is found the algorithm can often be accelerated.
 Although this slightly modified version is not complete anymore, in practice it is in some areas sufficient to know just one valid solution.
@@
 	\subsubsection{Depth-First Search with Optimization}
-	If in the constraint programming an objective is defined which should be either minimized or maximized, a modified version of the Depth-first search can be applied.
+	If in the constraint programming an objective is defined which should be either minimized or maximized, a modified version of the depth-first search can be applied.
 The only difference is that it evaluates the objective function for every found solution concerning the constraints and stores during the search only the best one.
@@
 
-	This algorithm is again a complete one, therefore it will always find the best solution if one exists.
-This algorithm does not inherit the bad memory characteristics from the Depth-first search as only one solution is stored, but the search is still exhaustive and can be very slow, depending on the size of the variable domains.
+	This algorithm is also a complete one.
+Hence, it will always find the best solution if one exists.
+This algorithm does not inherit the bad memory characteristics from the depth-first search as only one solution is stored, but the search is still exhaustive and can be very slow, depending on the size of the variable domains.
 
@@
 	A branch-and-bound search is an algorithm for optimization of an objective function under certain constraints and only discrete possible candidates for a solution.
-To fulfill the constraints the first step, branch-and-reduce is, applied, and the second one, bounding, is used to prune whole areas with non-optimal solutions.
+The difference to the depth-first search with optimization is the usage of a so-called relaxed optimum, the optimum without consideration of the required discreetness.
+
+	During a branch-and-bound search two steps are executed: branch-and-reduce and bounding. The first one partitions the search space and the second one prunes whole areas with non-optimal solutions.
 
@@
 
-	As the basic concept for finding solutions in Branch-and-bound is an enumeration of all possible candidates, it is necessary to separate the search space in a useful way, which is the task for the Branch-and-reduce step.
+	The basic concept for finding solutions in branch-and-bound is an enumeration of all possible candidates, therefore it is necessary to separate the search space in a useful way, which is the task for the branch-and-reduce step.
 Actually this step is very similar to the concept of branching, it also depends on a certain heuristic to divide the search space into two or even more disjoint parts.
@@
 This process can be stopped if the number of solutions is reduced to an amount, where it is possible to enumerate all of them and check if they are optimal.
-A possible division into sub problems can be seen in \reffig{fig:branchAndBound}, where for legibility Branch-and-reduce was applied only twice: The first execution split the problem into $P_0$ and the second split the other half into $P_1$ and $P_2$.
+A possible division into sub problems can be seen in \reffig{fig:branchAndBound}, where for legibility branch-and-reduce was applied only twice: The first execution split the problem into $P_0$ and $P'_0$, which was then split again into $P_1$ and $P_2$.
 
-	The next step after Branch-and-reduce is the Bounding, which is able to prune whole sub problems.
-In the following part I will consider the minimization of an objective function without a loss of generality.
-For the Bounding it is first necessary to initialize a upper bound with a value, for example with $\infty$.
+	The next step after branch-and-reduce is the bounding, which is able to prune whole sub problems.
+In the following part I will consider the minimization of an objective function without a loss of generality:
+	\begin{equation}
+		\max_x \{f(x)\} = -\min_x \{-f(x)\}
+	\end{equation}
+
+	For the bounding it is first necessary to initialize a upper bound with a value, for example with $\infty$.
 This upper bound is updated during the process every time a new optimal solution is found to this new optimum.
@@
 	\subsubsection{Limited Discrepancy Search}
-	The Limited Discrepancy search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg in \cite{limitedDiscrepancy}.
-It depends on some heuristics for the tree, which are determined through branching strategies.
+	The limited discrepancy search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg \cite{limitedDiscrepancy}.
+It depends on branching heuristics for the tree, which are determined through branching strategies.
 The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive.
 Second, heuristics work worse the farther away from the solution they are applied, therefore decisions close to the root node are likely to fail.
-Consequently, in a chronological backtrack search the worst decisions are the most expensive ones.
-The Limited Discrepancy search solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later.
+Consequently, in a depth-first search the worst decisions are the most expensive ones.
+The limited discrepancy search solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later.
 The improvement compared to a chronological backtrack was shown by experimental results and theoretically proven in \cite{limitedDiscrepancy}.
@@
 	\subsubsection{Backjumping}
-	A specific version of backjumping was already mentioned, the Limited Discrepancy search, where instead of jumping one node back the algorithm always jumps as far back to the root node as possible.
+	A specific version of backjumping was already mentioned, the limited discrepancy search, where instead of jumping one node back the algorithm always jumps as far back to the root node as possible.
 In general this is proven to be better than the chronological backtracking \cite{limitedDiscrepancy}, but for certain problems it may be possible to find better nodes to jump back.
@@
 	Branching always means to apply certain heuristics and these heuristics may fail or produce bad results for the first execution of the search algorithm.
-If it turns out that the search tree is build in a disadvantageous way a restart strategy may do the trick to improve the search.
+If it turns out that the search tree is build in a disadvantageous way a restart strategy may be appropriate to improve the search.
 For such a restart the heuristic is improved, based on the things the search has learned to be a not so good solution during the first or first few attempts to build a search tree.
@@
 	\section{Conclusion}
-	\emph{Gecode}, as representative for an implementation of constraint programming, provides everything necessary to describe and solve arbitrary combinatorial or optimization problems.
-The interface allows an adaption to the specific area through modifications of algorithms, restarts or the implementation of a custom search algorithm, based on the already provided ones.
+	\emph{Gecode}, as representative for an implementation of constraint programming, provides the necessary tools to describe and solve a wide range of combinatorial or optimization problems.
+The interface allows an adaption to the area of application through modifications of algorithms, restarts or the implementation of a custom search algorithm, based on the already provided ones.
 Therefore \emph{Gecode} can be used in various areas, starting from scheduling, over portfolio optimization up to technical problems like placement.

### 2014-05-11 paper.tex
corrections

@@
 The root node is also the first one with a branch, where the search space is split up into disjoint sets.
-This approach is similar to the divide-and-conquer technique as it reduces a complex problem into two or more parts with less complexity.
+This approach is similar to the divide-and-conquer technique, as it reduces a complex problem into two or more parts with less complexity.
 The splitting, which adds additional constraints on the variables, is then recursively done with different heuristics to partition the search space into smaller parts.
@@
 	Constraints in \emph{Gecode} are evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the sub trees.
-The problem here is to decide in which order the constraints should be considered, and it may be even necessary to evaluate some constraints more often.
+The problem here is to decide in which order the constraints should be considered, and it may even be necessary to evaluate some constraints more often.
 A short example will illustrate this \cite[p.~575]{handbookCP}: Let there be two constraints
@@
 		\item Enumeration: The domain for $x \in D = {x_1, x_2, \dots}$ is split up into $|D|$-branches where for every branch $i$ one additional constraint $c:\ x = x_i$ is posted.
-		\item Binary choice points: One value $x_i$ of the domain $D$ of the variable $x$ is selected and the domain is split up into two sub trees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted.
+		\item Binary choice points: One value $x_i$ from the domain $D$ of the variable $x$ is selected and the domain is split up into two sub trees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted.
 This results into something similar to a binary tree at this point.
@@
 	As for variables with a float domain the enumeration and binary choice points are not feasible, the last one, domain splitting, is typically used.
-As termination criteria for a float it is often necessary to define a certain epsilon $\epsilon > 0$ which describes the interval length, at which the solver states to have found a solution.
+As termination criteria for a float it is necessary to define a certain epsilon $\epsilon > 0$ which describes the interval length, at which the solver states to have found a solution.
 
@@
 	\subsubsection{Depth-First Search}
-	The depth-first search was historical the first one which was developed.
+	The depth-first search was historical the first one which was developed, therefore it is a good point to start explaining search algorithms.
 The idea behind this search algorithm is that every possible combination of variables is examined.
@@
 If a new possible solution is found which is above this upper bound the algorithm can prune it, as only the minimum solution is of interest.
-The actually pruning of sub problems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
+
+	The actual pruning of sub problems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
 If this relaxed optimum for a certain sub problem is greater than the current upper bound it is possible to prune this complete sub problem, as it can not contain a better solution.
@@
 	\subsubsection{Limited Discrepancy Search}
-	The limited discrepancy search falls into the category of Best-first searches and was first mentioned by Harvey and Ginsberg \cite{limitedDiscrepancy}.
+	The limited discrepancy search falls into the category of best-first searches and was first mentioned by Harvey and Ginsberg \cite{limitedDiscrepancy}.
 It depends on branching heuristics for the tree, which are determined through branching strategies.
@@
 	A specific version of backjumping was already mentioned, the limited discrepancy search, where instead of jumping one node back the algorithm always jumps as far back to the root node as possible.
-In general this is proven to be better than the chronological backtracking \cite{limitedDiscrepancy}, but for certain problems it may be possible to find better nodes to jump back.
+In general this is proven to be better than the chronological backtracking \cite{limitedDiscrepancy}.
+However, for certain problems it may be possible to find better nodes to jump back.
 A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the dead end.

### 2014-05-12 paper.tex
corrections of punctuation

@@
 	\subsection{Variable Domains}
-	In constraint programming, a variable is connected to a set of possible values, the so called domain.
+	In constraint programming a variable is connected to a set of possible values, the so called domain.
 Depending on the domain, a variable is called an integer, Boolean or floating variable.
@@
 
-	In \emph{Gecode}, it is possible to define custom strategies. However, the user is supported at this point with some standard algorithms which he can use to build custom ones:
+	In \emph{Gecode} it is possible to define custom strategies. However, the user is supported at this point with some standard algorithms which he can use to build custom ones:
 	\begin{itemize}
@@
 
-	In the following part, I will discuss the depth-first search, the branch-and-bound search, the limited discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
+	In the following part I will discuss the depth-first search, the branch-and-bound search, the limited discrepancy search as candidate for an improved backjumping-strategy, backjumping strategies in general and restart strategies.
 
@@
 The idea behind this search algorithm is that every possible combination of variables is examined.
-In its basic form, it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance branch-and-reduce.
+In its basic form it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance branch-and-reduce.
 Sometimes the depth-first search is also referred to as chronological backtrack search.
@@
 	This search is a so-called complete algorithm as it can be proven that all possible solutions will be found \cite[p.~85]{handbookCP}.
-Because of this characteristic, it is very useful to prove for example that for a certain problem no solution exists.
+Because of this characteristic it is very useful to prove for example that for a certain problem no solution exists.
 The main drawback is the execution time, as it does an exhaustive search.

### 2014-05-13 paper.tex
corrections Michael

@@
 \begin{document}
-	\title{Constraint Programming - Inside Gecode}
+	\title{Constraint Programming Algorithms used in Gecode}
 	\author{Benedikt~Schmidt}
@@
 
-	\begin{figure}
+	\begin{figure}[b]
 		\center
@@
 
-	In the following I will explain the necessary steps to describe and solve a problem with constraint programming: Variables, constraints, search trees, constraint propagation and search algorithms.
+	In the following I will explain the necessary steps to describe and solve a problem with constraint programming: variables, constraints, search trees, constraint propagation and search algorithms.
 As example of an implementation I will refer to \emph{Gecode} \cite{gecode}.
@@
 	\begin{itemize}
-		\item Integer: finite or infinite sets like $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$, \dots
+		\item Integer: finite or infinite, e.g. $x \in \mathbb{N}$, $x \in \{-5, 4, 10\}$
 		\item Bool: $x \in \{\text{true}, \text{false}\}$
-		\item Float: intervals or unions of intervals like $x \in [-4, 3]$, $x \in [-10, 8] \cup [10, 12]$, \dots
+		\item Float: intervals or unions of intervals, e.g. $x \in [-4, 3]$, $x \in [-10, 8] \cup [10, 12]$
 	\end{itemize}
@@
 	\subsection{Constraint Types}
-	Which statements are possible to make and how comfortable it is to do so varies from the programming language in which the problem is implemented.
+	The choice of constraint types and how comfortable it is to specify them varies from the programming language in which the problem is implemented.
 For sure, there is also a difference in the interfaces of the libraries for constraint programming.
@@
 Some of these constraints may acquire special attention during the search for solutions to improve the runtime of the search.
-For this reason a lot of work has been done for example on the alldifferent constraint \refeq{eq:alldifferent} \cite{allDifferent}, which would be such a combined constraint.
+For this reason a lot of work has been done for example on the alldifferent constraint \cite{allDifferent}
 	\begin{equation}
@@
 	\end{equation}
+	which is an example for a combined constraint.
 
@@
 The root node is also the first one with a branch, where the search space is split up into disjoint sets.
-This approach is similar to the divide-and-conquer technique, as it reduces a complex problem into two or more parts with less complexity.
+This approach is based on the same idea like the divide-and-conquer technique \cite[p.~175]{artOfComputerProgramming}, as it reduces a complex problem into two or more parts with less complexity.
 The splitting, which adds additional constraints on the variables, is then recursively done with different heuristics to partition the search space into smaller parts.
@@
 The term propagation refers in this area to the reduction of choices for variables which are infeasible, considering other constraints and affected variable domains.
-As the size of the variable domain is related to the possibilities which have to be checked by a search through additional subtrees, the constraint propagation is able to prune complete infeasible sub trees.
+As the size of the variable domain is related to the possibilities which have to be checked by a search through additional subtrees, the constraint propagation is able to prune complete infeasible subtrees.
 Therefore, the whole search process can be accelerated through constraint propagation.
@@
 
-	To illustrate this effect consider a constraint problem based on two variables $x \in \{1, 5, 6\}$ and $y \in \{1, 2\}$ and one constraint based on them.
+	To illustrate this effect consider a problem based on two variables $x \in \{1, 5, 6\}$ and $y \in \{1, 2\}$ and the constraint
 	\begin{equation}
@@
 By inspection one sees that the right parts of this tree can be already pruned at the root by propagation of constraint $c_1$.
-Therefore, it is not necessary to create and evaluate these sub trees.
+Therefore, it is not necessary to create and evaluate these subtrees.
 With constraint propagation fewer branches have to be made, like it can be seen in \reffig{fig:smallTree}.
@@
 
-	Constraints in \emph{Gecode} are evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the sub trees.
+	Constraints in \emph{Gecode} are evaluated ahead of a further growth of a subtree, as it may be possible to prune some of the subtrees.
 The problem here is to decide in which order the constraints should be considered, and it may even be necessary to evaluate some constraints more often.
@@
 	To implement this repeated evaluation of constraints in \emph{Gecode} all constraints subscribe to the domains of the variables which are part of them.
+A more detailed explanation of this observer pattern can be found in \cite{designPatterns}.
 With this method, the constraints get notified whenever a variable domain is reduced and they can signal the system that they need to be evaluated once again.
@@
 The question is answered by the selected branching strategy.
-No matter which one is selected, they all have in common that they create additional, temporary constraints, which are only applied to the specific sub trees.
+No matter which one is selected, they all have in common that they create additional, temporary constraints, which are only applied to the specific subtrees.
 This process is called \emph{posting} a constraint on a branch.
 
-	The branching strategy depends on the variable domain: For integer variables there are three common strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the sub trees.
+	The branching strategy depends on the variable domain: For integer variables there are three common strategies \cite[p.~87]{handbookCP} which differ in how the domain is distributed to the subtrees.
 	\begin{itemize}
 		\item Enumeration: The domain for $x \in D = {x_1, x_2, \dots}$ is split up into $|D|$-branches where for every branch $i$ one additional constraint $c:\ x = x_i$ is posted.
-		\item Binary choice points: One value $x_i$ from the domain $D$ of the variable $x$ is selected and the domain is split up into two sub trees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted.
+		\item Binary choice points: One value $x_i$ from the domain $D$ of the variable $x$ is selected and the domain is split up into two subtrees: On one the constraint $x = x_i$ and on the other one the constraint $x \ne x_i$ is posted.
 This results into something similar to a binary tree at this point.
@@
 	\subsubsection{Depth-First Search}
-	The depth-first search was historical the first one which was developed, therefore it is a good point to start explaining search algorithms.
 	The idea behind this search algorithm is that every possible combination of variables is examined.
-In its basic form it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance branch-and-reduce.
+In its basic form it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance branch-and-bound.
 Sometimes the depth-first search is also referred to as chronological backtrack search.
@@
 	The starting point for the depth-first search is the root node, from where on the algorithm is called recursively.
-At every node, mainly two steps are executed: First, a constraint propagation to reduce the variable domains and second a branching, which divides the search space into two disjoint sub spaces, on which the same algorithm is applied again.
+At every node, mainly two steps are executed: First, a constraint propagation to reduce the variable domains and second a branching, which divides the search space into two disjoint subspaces, on which the same algorithm is applied again.
 The search stops in a leaf of the search tree if either no branching is possible and therefore a valid solution is found, or one or more variable domains are emptied, which indicates that in this leave no possible solution exists.
@@
 
-	By stopping the search after the first feasible solution is found the algorithm can often be accelerated.
+	By stopping the search after the first feasible solution is found the runtime of the search can be improved.
 Although this slightly modified version is not complete anymore, in practice it is in some areas sufficient to know just one valid solution.
@@
 The only difference is that it evaluates the objective function for every found solution concerning the constraints and stores during the search only the best one.
-To accelerate the search it is possible to add an additional constraint which defines that the new objective must better than the best so far.
-Through this enhancement even more sub trees are pruned and the search is faster.
+To accelerate the search it is possible to add an additional constraint which requires that the new value for the objective must be better than the best achieved so far.
+Through this enhancement even more subtrees are pruned and the search is faster.
 
-	This algorithm is also a complete one.
-Hence, it will always find the best solution if one exists.
+	This algorithm is also complete.
+Hence, it will always find the best solution, if one exists.
 This algorithm does not inherit the bad memory characteristics from the depth-first search as only one solution is stored, but the search is still exhaustive and can be very slow, depending on the size of the variable domains.
@@
 		\end{tikzpicture}	
-		\caption{Splitting of a search space into three sub problems and the possible candidates for solutions}
+		\caption{Splitting of a search space into three subproblems and the possible candidates for solutions}
 		\label{fig:branchAndBound}
@@
 
-	The basic concept for finding solutions in branch-and-bound is an enumeration of all possible candidates, therefore it is necessary to separate the search space in a useful way, which is the task for the branch-and-reduce step.
-Actually this step is very similar to the concept of branching, it also depends on a certain heuristic to divide the search space into two or even more disjoint parts.
+	The basic concept for finding solutions in branch-and-bound is an enumeration of all possible candidates.
+Therefore, it is necessary to separate the search space in an appropriate way, which is the task for the branch-and-reduce step.
+This step is very similar to the concept of branching, it also depends on a certain heuristic to divide the search space into two or even more disjoint parts.
 This heuristic can either be selected by the user or be a general one, like splitting the biggest variable domain into two equal sized parts.
-If the basic search space contains a big amount of possible candidate solutions one reduction may not be enough and therefore this step is applied recursively to the newly create sub problems.
+If the basic search space contains a big amount of possible candidate solutions one reduction may not be enough and therefore this step is applied recursively to the newly create subproblems.
 This process can be stopped if the number of solutions is reduced to an amount, where it is possible to enumerate all of them and check if they are optimal.
-A possible division into sub problems can be seen in \reffig{fig:branchAndBound}, where for legibility branch-and-reduce was applied only twice: The first execution split the problem into $P_0$ and $P'_0$, which was then split again into $P_1$ and $P_2$.
+A possible division into subproblems can be seen in \reffig{fig:branchAndBound}, where for legibility branch-and-reduce was applied only twice: The first execution split the problem into $P_0$ and $P'_0$, which was then split again into $P_1$ and $P_2$.
 
-	The next step after branch-and-reduce is the bounding, which is able to prune whole sub problems.
+	The next step after branch-and-reduce is the bounding, which is able to prune whole subproblems.
 In the following part I will consider the minimization of an objective function without a loss of generality:
@@
 
-	The actual pruning of sub problems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
-If this relaxed optimum for a certain sub problem is greater than the current upper bound it is possible to prune this complete sub problem, as it can not contain a better solution.
+	The actual pruning of subproblems is based on the relaxed optimum, which is the optimum of the problem without consideration that only discrete values are allowed.
+If this relaxed optimum for a certain subproblem is greater than the current upper bound it is possible to prune this complete subproblem, as it can not contain a better solution.
 
-	Consider for this \reffig{fig:branchAndBound}, where already a certain optimum $f$ was found in sub problem $P_1$.
+	Consider for this \reffig{fig:branchAndBound}, where already a certain optimum $f$ was found in subproblem $P_1$.
 If now the relaxed optimum $f_{P0}$ in $P_0$ and $f_{P2}$ in $P_2$ is worse than this optimum
@@
 	\end{equation}
-	the whole sub problems $P_{0}$ and $P_{2}$ can be pruned and it is not necessary to enumerate the solution candidates in them (or to split them further up).
+	the whole subproblems $P_{0}$ and $P_{2}$ can be pruned and it is not necessary to enumerate the solution candidates in them (or to split them further up).
 
@@
 The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive.
-Second, heuristics work worse the farther away from the solution they are applied, therefore decisions close to the root node are likely to fail.
+Second, heuristics work worse the farther away from the solution they are applied.
+Therefore, decisions close to the root node are likely to fail.
 Consequently, in a depth-first search the worst decisions are the most expensive ones.
@@
 For such a restart the heuristic is improved, based on the things the search has learned to be a not so good solution during the first or first few attempts to build a search tree.
-	For practical application several variants of restart strategies were proposed \cite[p.~113]{handbookCP}
+	For practical application several variants of restart strategies were proposed \cite[p.~113]{handbookCP}.
 
@@
 The interface allows an adaption to the area of application through modifications of algorithms, restarts or the implementation of a custom search algorithm, based on the already provided ones.
-Therefore \emph{Gecode} can be used in various areas, starting from scheduling, over portfolio optimization up to technical problems like placement.
+Therefore, \emph{Gecode} can be used in various areas, starting from scheduling, over portfolio optimization up to technical problems like placement.
 
@@
 		\bibitem{handbookCP}
-		F.~Rossi, P.~van~Beek and T.~Walsh, \emph{Handbook of Constraint Programming}, Amsterdam, The Netherlands: Elsevier, 2006, ISBN: 978-0-444-52726-4
+		F.~Rossi, P.~van~Beek and T.~Walsh, \emph{Handbook of Constraint Programming}, Amsterdam, The Netherlands: Elsevier, 2006, ISBN-13: 978-0-444-52726-4
 		\bibitem{allDifferent}
@@
 		\bibitem{trendsInCP}
-		F.~Benhamou, N.~Jussien, B.~A.~O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN: 978-1-905209-97-2
+		F.~Benhamou, N.~Jussien, B.~A.~O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN-13: 978-1-905209-97-2
 		\bibitem{gecode}
@@
 		\bibitem{linearProgramming}
-		L.~Cooper and D.~Steinberg, \emph{Methods and Applications of Linear Programming}, W.~B.~Saunder~Company, 1974, ISBN: 0-7216-2694-7
+		L.~Cooper and D.~Steinberg, \emph{Methods and Applications of Linear Programming}, W.~B.~Saunder~Company, 1974, ISBN-10: 0-7216-2694-7
+		\bibitem{artOfComputerProgramming}
+		D.~E.~Knuth, \emph{The Art of Computer Programming: Volume 3, Sorting and Searching}, second edition, Addison-Wesley, 1998, ISBN-10: 0201896850, ISBN-13: 9780201896855
+		\bibitem{designPatterns}
+		E.~Gamma, R.~Helm, R.~Johnson, J.~Vlissides, \emph{Design Patterns. Elements of Reusable Object-Oriented Software}, first edition, Prentice Hall, 1994, ISBN-10: 0201633612
 	\end{thebibliography}

### 2014-05-19 paper.tex
added objective functions

@@
 
-	In the following I will explain the necessary steps to describe and solve a problem with constraint programming: variables, constraints, search trees, constraint propagation and search algorithms.
+	In the following I will explain the necessary steps to describe and solve a problem with constraint programming: variables, constraints, objective functions, search trees, constraint propagation and search algorithms.
 As example of an implementation I will refer to \emph{Gecode} \cite{gecode}.
@@
 	which is an example for a combined constraint.
+
+	\subsection{Objective Functions}
+	A constraint problem can have, additional to constraints, an objective function.
+As in optimization only one objective function can be optimized at a time, a problem in constraint programming must have either one or no objective function.
+To solve problems with an objective function it is necessary to select a search algorithm which is capable of optimization.
 

### 2014-05-19 paper.tex
improved figure for branch-and-bound
moved figure
corrections Michael

@@
 	\label{fig:smallTree}
+	\end{figure}\begin{figure}
+		\center
+		\subfloat[First step]
+		{
+			\begin{tikzpicture}
+				\coordinate (a) at (1, 0);
+				\coordinate (b) at (3, 0);
+				\coordinate (c) at (3.5, 2);
+				\coordinate (d) at (2, 4);
+				\coordinate (e) at (0.5, 1);
+				\coordinate (f) at (2, 0);
+				\coordinate (g) at (intersection of e--c and f--d);
+
+				\begin{scope}
+					\clip (a) -- (b) -- (c) -- (d) -- (e) -- (a);
+					\foreach \x in {-2,-1.5,...,5}
+					{
+		      			\foreach \y in {0,0.5,...,6}
+		      			{
+		        			\node[draw,circle,inner sep=0.2pt,fill,gray] at (\x,\y) {};
+		      			}
+		    		}
+	    		\end{scope}
+
+				\draw[very thick] (a) -- (b) -- (c) -- (d) -- (e) -- (a);
+				\draw (a) -- (b) -- (c) -- (e) -- (a);
+				\draw[fill=blue, fill opacity=0.2] (c) -- (d) -- (e) -- (c);
+				\draw[fill=green, fill opacity=0.2] (a) -- (b) -- (c) -- (e) -- (a);
+				\node at (2.25, 2.75) {$P_0$};
+				\node at (1.75, 0.75) {$P_0'$};
+			\end{tikzpicture}	
+		}
+		\subfloat[Second step]
+		{		
+			\begin{tikzpicture}
+				\coordinate (a) at (1, 0);
+				\coordinate (b) at (3, 0);
+				\coordinate (c) at (3.5, 2);
+				\coordinate (d) at (2, 4);
+				\coordinate (e) at (0.5, 1);
+				\coordinate (f) at (2, 0);
+				\coordinate (g) at (intersection of e--c and f--d);
+
+				\begin{scope}
+					\clip (a) -- (b) -- (c) -- (d) -- (e) -- (a);
+					\foreach \x in {-2,-1.5,...,5}
+					{
+		      			\foreach \y in {0,0.5,...,6}
+		      			{
+		        			\node[draw,circle,inner sep=0.2pt,fill,gray] at (\x,\y) {};
+		      			}
+		    		}
+	    		\end{scope}
+
+				\draw[very thick] (a) -- (b) -- (c) -- (d) -- (e) -- (a);
+				\draw (a) -- (b) -- (c) -- (e) -- (a);
+				\draw (f) -- (g) -- (c) -- (b) -- (f);
+				\draw[fill=blue, fill opacity=0.2] (c) -- (d) -- (e) -- (c);
+				\draw[fill=green, fill opacity=0.2] (a) -- (f) -- (g) -- (e) -- (a);
+				\draw[fill=brown, fill opacity=0.2] (f) -- (g) -- (c) -- (b) -- (f);
+				\node at (2.25, 2.75) {$P_0$};
+				\node at (1.25, 0.75) {$P_1$};
+				\node at (2.75, 1.25) {$P_2$};
+			\end{tikzpicture}	
+		}
+		\caption{Splitting of a search space into three subproblems and the possible candidates for solutions}
+		\label{fig:branchAndBound}
 	\end{figure}
@@
 
-	\begin{figure}
-		\center
-		\begin{tikzpicture}
-			\coordinate (a) at (0, 0);
-			\coordinate (b) at (4, 0);
-			\coordinate (c) at (5, 3);
-			\coordinate (d) at (3, 6);
-			\coordinate (e) at (-1, 2);
-			\coordinate (f) at (2, 0);
-			\coordinate (g) at (intersection of e--c and f--d);
-
-			\begin{scope}
-				\clip (a) -- (b) -- (c) -- (d) -- (e) -- (a);
-				\foreach \x in {-2,-1.5,...,5}
-				{
-	      			\foreach \y in {0,0.5,...,6}
-	      			{
-	        			\node[draw,circle,inner sep=0.2pt,fill,gray] at (\x,\y) {};
-	      			}
-	    		}
-    		\end{scope}
-
-			\draw[very thick] (a) -- (b) -- (c) -- (d) -- (e) -- (a);
-			\draw (a) -- (b) -- (c) -- (e) -- (a);
-			\draw (f) -- (g) -- (c) -- (b) -- (f);
-			\draw[fill=blue, fill opacity=0.2] (c) -- (d) -- (e) -- (c);
-			\draw[fill=green, fill opacity=0.2] (a) -- (f) -- (g) -- (e) -- (a);
-			\draw[fill=brown, fill opacity=0.2] (f) -- (g) -- (c) -- (b) -- (f);
-			\node at (2.75, 3.75) {$P_0$};
-			\node at (1.25, 1.25) {$P_1$};
-			\node at (3.75, 1.75) {$P_2$};
-		\end{tikzpicture}	
-		\caption{Splitting of a search space into three subproblems and the possible candidates for solutions}
-		\label{fig:branchAndBound}
-	\end{figure}
-
 	The basic concept for finding solutions in branch-and-bound is an enumeration of all possible candidates.
@@
 This step is very similar to the concept of branching, it also depends on a certain heuristic to divide the search space into two or even more disjoint parts.
-This heuristic can either be selected by the user or be a general one, like splitting the biggest variable domain into two equal sized parts.
+The heuristic can either be selected by the user or be a general one, like splitting the biggest variable domain into two equal sized parts.
 If the basic search space contains a big amount of possible candidate solutions one reduction may not be enough and therefore this step is applied recursively to the newly create subproblems.
@@
 It depends on branching heuristics for the tree, which are determined through branching strategies.
-The basic concept is that during a chronological backtrack search a wrong turn close to the root is first of all very expensive.
+The basic concept is that during a depth-first search a wrong turn close to the root is first of all very expensive.
 Second, heuristics work worse the farther away from the solution they are applied.

### 2014-05-19 paper.tex
corrections Michael

@@
 If it turns out that the search tree is build in a disadvantageous way a restart strategy may be appropriate to improve the search.
-For such a restart the heuristic is improved, based on the things the search has learned to be a not so good solution during the first or first few attempts to build a search tree.
+For such a restart the heuristic is improved, based on the things the search has learned to be a not so good solution during the first few attempts to build a search tree.
 	For practical application several variants of restart strategies were proposed \cite[p.~113]{handbookCP}.

### 2014-05-19 paper.tex
reduced to four pages

@@
 A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the dead end.
-How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is done wrong the backjumping can even be less efficient than chronological backtracking \cite[p.~100]{handbookCP}.
+How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is done wrong the backjumping can even be less efficient than a depth-first search \cite[p.~100]{handbookCP}.
 

### 2014-06-17 paper.tex
fixed title of seminar
improvements

@@
 	\title{Constraint Programming Algorithms used in Gecode}
-	\author{Benedikt~Schmidt}
-	\markboth{Advanced Seminar for Security in Information Technology, Summer Term 2014}%
+	\author{Benedikt~Schmidt <benediktibk@aon.at>}
+	\markboth{Advanced Seminar for VLSI Design Methods, Summer Term 2014}%
 	{Benedikt Schmidt: Constraint Programming - Inside \emph{Gecode}}
@@
 
-	\begin{figure}[b]
+	\begin{figure}[h]
 		\center
@@
 	In the following I will explain the necessary steps to describe and solve a problem with constraint programming: variables, constraints, objective functions, search trees, constraint propagation and search algorithms.
-As example of an implementation I will refer to \emph{Gecode} \cite{gecode}.
+As example of an implementation I will refer to \emph{Gecode}\footnote{http://www.gecode.org/}.
 
@@
 	\end{itemize}
-	This distinction is fundamental as especially for floating variables a combinatorial approach to constraint programming is impossible.
+	This distinction is fundamental as especially for floating variables a combinatorial approach is impossible.
 
@@
 For sure, there is also a difference in the interfaces of the libraries for constraint programming.
-Typical constraints are
+Typical constraints are:
 	\begin{itemize}
@@
 		F.~Benhamou, N.~Jussien, B.~A.~O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN-13: 978-1-905209-97-2
-		\bibitem{gecode}
-		http://www.gecode.org/
 		\bibitem{programmingGecode}

### 2014-06-17 paper.tex
added abbreviations

@@
 
-	\subsubsection{Depth-First Search}
+	\subsubsection{Depth-First Search (DFS)}
 	The idea behind this search algorithm is that every possible combination of variables is examined.
 In its basic form it can not be applied to problems with infinite domains; for problems with such variable domains it is necessary to use for instance branch-and-bound.
-Sometimes the depth-first search is also referred to as chronological backtrack search.
+Sometimes the DFS is also referred to as chronological backtrack search.
 
-	The starting point for the depth-first search is the root node, from where on the algorithm is called recursively.
+	The starting point for the DFS is the root node, from where on the algorithm is called recursively.
 At every node, mainly two steps are executed: First, a constraint propagation to reduce the variable domains and second a branching, which divides the search space into two disjoint subspaces, on which the same algorithm is applied again.
@@
 	\subsubsection{Depth-First Search with Optimization}
-	If in the constraint programming an objective is defined which should be either minimized or maximized, a modified version of the depth-first search can be applied.
+	If in the constraint programming an objective is defined which should be either minimized or maximized, a modified version of the DFS can be applied.
 The only difference is that it evaluates the objective function for every found solution concerning the constraints and stores during the search only the best one.
@@
 Hence, it will always find the best solution, if one exists.
-This algorithm does not inherit the bad memory characteristics from the depth-first search as only one solution is stored, but the search is still exhaustive and can be very slow, depending on the size of the variable domains.
+This algorithm does not inherit the bad memory characteristics from the DFS as only one solution is stored, but the search is still exhaustive and can be very slow, depending on the size of the variable domains.
 
@@
 	A branch-and-bound search is an algorithm for optimization of an objective function under certain constraints and only discrete possible candidates for a solution.
-The difference to the depth-first search with optimization is the usage of a so-called relaxed optimum, the optimum without consideration of the required discreetness.
+The difference to the DFS with optimization is the usage of a so-called relaxed optimum, the optimum without consideration of the required discreetness.
 
@@
 
-	\subsubsection{Limited Discrepancy Search}
-	The limited discrepancy search falls into the category of best-first searches and was first mentioned by Harvey and Ginsberg \cite{limitedDiscrepancy}.
+	\subsubsection{Limited Discrepancy Search (LDS)}
+	The LDS falls into the category of best-first searches and was first mentioned by Harvey and Ginsberg \cite{limitedDiscrepancy}.
 It depends on branching heuristics for the tree, which are determined through branching strategies.
-The basic concept is that during a depth-first search a wrong turn close to the root is first of all very expensive.
+The basic concept is that during a DFS a wrong turn close to the root is first of all very expensive.
 Second, heuristics work worse the farther away from the solution they are applied.
 Therefore, decisions close to the root node are likely to fail.
-Consequently, in a depth-first search the worst decisions are the most expensive ones.
-The limited discrepancy search solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later.
+Consequently, in a DFS the worst decisions are the most expensive ones.
+The LDS solves exactly this problem as it suggests to change rather the decisions made close to the root than the ones the search made later.
 The improvement compared to a chronological backtrack was shown by experimental results and theoretically proven in \cite{limitedDiscrepancy}.
@@
 A jump back in the search tree means to detect a so-called nogood decision, which caused the violation of a constraint in the dead end.
-How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is done wrong the backjumping can even be less efficient than a depth-first search \cite[p.~100]{handbookCP}.
+How these nogoods must be selected and stored to improve the search depends heavily on the specific field and if it is done wrong the backjumping can even be less efficient than a DFS \cite[p.~100]{handbookCP}.
 
